---
title: "Fast followers?/Speed contagion: assessing the impact of Montreal F1 Grand Prix on high-speed ticketing rates (2000-2022)"
description: |
  Step 3. Check analytical strategies (part 2)
lang: en
date: last-modified #date-format: "D [d]e MMM, YYYY"
bibliography: [_lit/refs.bib]
csl: "_lit/american-medical-association.csl" # (Opcional: estilo de citas, p.ej., APA)
author: "Andrés González Santa Cruz"
format: 
  html:
    css: [_lib/styles.scss]
    code-fold: true
    embed-resources: true
    fig-cap-location: top
lightbox: auto
toc: true
toc-depth: 5
toc-location: left
toc-float: true
editor: source
---

<style>
.scrollable-content {
  max-height: 350px;
  overflow-y: auto;
}
</style>
<style>
pre.scrollable-code {
  max-height: 350px;
  overflow-y: auto;
}
</style>

```{r}
#| message: true
#| include: true
#| warning: true

# remove objects and memory
rm(list=ls());gc()

#remove images
while(!dev.cur())dev.off()
cat("\014")
load(paste0(getwd(),"/_data/step2.RData"))
```

# Load libraries and data

<div class="scrollable-content">
```{r}
#| label: setup
#| message: true
#| warning: true
#| class-output: center-table
#| include: true
#| results: "hold"

#borrar caché
#system("fc-cache -f -v")

#check R version
if(Sys.info()["sysname"]=="Windows"){
if (getRversion() != "4.4.1") { stop("Requiere versión de R 4.4.1. Actual: ", getRversion()) }
}
if(Sys.info()["sysname"]=="Linux"){
if (getRversion() != "4.4.1") { stop("Requiere versión de R 4.4.1. Actual: ", getRversion()) }
}
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:

# install.packages(c("dplyr", #for data
#                    "tidyr", #for data
#                    "lubridate",  #for dates
#                    "openxlsx", #for excel files
#                    "rio", #for importing and exporting data
#                    "purrr", #for iterating in databases
#                    "devtools", #for external packages
#                    "DiagrammeR", #para visualizar DAG
#                    "dagitty",
#                    "ggdag",
#                    "ggplot2" #for graphics
#                    "kableExtra", #pretty tables
#                    "quarto", #for documents
#                    "geosphere" #for coordinates and classifying
#                    "geepack", #for regression
#                    "glmmTMB", #For GLMMs
#                    "DHARMa", #For residual diagnostics
#                    "car", #For hypothesis testing
#                    "brms", #Bayesian model
#                    "bayesplot",
#                    "loo",
#                    "Synth", #for synthetic control method
#                    "weathercan",#for weather data
#                    "sandwich", #cluster robust intervals
#                    "emmeans", #for predictions
#                    "gnm", #Conditional gaussian models
#                    "splines", #nonlinearity 
#                    "geeM", #negative binomial and more flexible GEE models
#                    "PanelMatch", #Matching technique with panel data
#                    "scpi" #control sintético
#                    "nixtlar", #for time series analysis and prediction
#                    "CausalImpact" #for time series causal impact
#                    "forecast" #for time series analysis prediction and decomposition
#                    ))
library(dplyr); library(lubridate); library(tidyverse); library(openxlsx); library(rio); library(purrr); library(dagitty); library(ggdag); library(kableExtra); library(geosphere); library(geepack); library(lme4); library(glmmTMB); library(DHARMa); library(car); library(brms); library(bayesplot); library(loo); library(Synth);  library(weathercan); library(sandwich); library(emmeans); library(gnm); library(splines); library(geeM); library(plm); library(PanelMatch); library(scpi); library(fect); library(nixtlar); library(CausalImpact); library(forecast)

#special repository indicated or the package
if(!require(weathercan)){
   install.packages("weathercan", 
                  repos = c("https://ropensci.r-universe.dev", "https://cloud.r-project.org")); library(weathercan)
  }

#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_

if(!require(bpmn)){devtools::install_github("bergant/bpmn")}

#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
sum_dates <- function(x){
 
  cbind.data.frame(
	min= as.Date(min(unclass(as.Date(x)), na.rm=T), origin = "1970-01-01"),
	p001= as.Date(quantile(unclass(as.Date(x)), .001, na.rm=T), origin = "1970-01-01"),
	p005= as.Date(quantile(unclass(as.Date(x)), .005, na.rm=T), origin = "1970-01-01"),
	p025= as.Date(quantile(unclass(as.Date(x)), .025, na.rm=T), origin = "1970-01-01"),
	p25= as.Date(quantile(unclass(as.Date(x)), .25, na.rm=T), origin = "1970-01-01"),
	p50= as.Date(quantile(unclass(as.Date(x)), .5, na.rm=T), origin = "1970-01-01"),
	p75= as.Date(quantile(unclass(as.Date(x)), .75, na.rm=T), origin = "1970-01-01"),
	p975= as.Date(quantile(unclass(as.Date(x)), .975, na.rm=T), origin = "1970-01-01"),
	p995= as.Date(quantile(unclass(as.Date(x)), .995, na.rm=T), origin = "1970-01-01"),
	p999= as.Date(quantile(unclass(as.Date(x)), .999, na.rm=T), origin = "1970-01-01"),
	max= as.Date(max(unclass(as.Date(x)), na.rm=T), origin = "1970-01-01")
  )
}
smd_bin <- function(x,y){
  z <- x*(1-x)
  t <- y*(1-y)
  k <- sum(z,t)
  l <- k/2
  
  return((x-y)/sqrt(l))
  
}

theme_custom_sjplot2 <- function(base_size = 12, base_family = "") {
  theme_minimal(base_size = base_size, base_family = base_family) +
    theme(
      # Text elements
      text = element_text(size = base_size, family = base_family),
      plot.title = element_text(face = "bold", hjust = 0.5, size = base_size * 1.2),
      plot.subtitle = element_text(hjust = 0.5, margin = margin(b = 10)),
      axis.title = element_text(size = base_size, face = "bold"),
      axis.text = element_text(size = base_size * 0.8),
      axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.5),
      axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5),
      axis.title.x = element_text(margin = margin(t = 10)),
      axis.title.y = element_text(margin = margin(r = 10)),
      
      # Plot layout
      plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
      panel.grid.major = element_line(color = "grey80"),
      panel.grid.minor = element_blank(),
      legend.position = "right",
      legend.text = element_text(size = base_size * 0.8),
      legend.title = element_text(size = base_size, face = "bold"),
      legend.background = element_rect(fill = "white", colour = NA),
      legend.box.background = element_rect(colour = "grey80", linetype = "solid"),
      legend.key = element_rect(fill = "white", colour = "white")
    )
}

#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
num_cores <- parallel::detectCores() -1
data.table::setDTthreads(threads = num_cores)#restore_after_fork = NULL, throttle = NULL)

#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#CONFIG #######################################################################
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_

options(scipen=2) #display numbers rather scientific number


nixtlar::nixtla_client_setup(api_key = readLines(paste0(gsub("f1/","", getwd() ),"/key_alt.txt"))[[1]])

```


</div>

### Matrix-completion method

```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: false
#| label: gsynth-mc
#| eval: true
#| result: "hold"
#| fig.cap: "Prediciton on the treated (adjusting on min max temperature and lagged precipitiation (2 days) and lagged rates"
#| fig-width: 9.5
#| fig-height: 5.5
#| fig-dpi: 600
#| results: "hold"
df <- subset(
  collisions_weather_corr_rect_synth,
  yday_corr <= (max_time - 7)            # mismo recorte temporal que en countSynth
)
df$treat_unit <- df$tr_contrl_corr
df$post_treat <- ifelse(df$yday_corr >= 35, 1, 0)
df$D          <- df$tr_contrl_corr * df$post_treat   # interacción exposición x tiempo

df$log_rate <- log(df$sum_velocidad_lead2) - log(df$exp_lic)     # log(count) – log(exposure)

#table(collisions_weather_corr_rect_synth$id_lab,collisions_weather_corr_rect_synth$tr_contrl_corr)

library(fect)
set.seed(2125)

road.main.gsynth <- fect::fect(
  formula = rate_lead2_lic ~ D + mean_min_temp_mean_lin + mean_max_temp_mean_lin +#rate_lead2_lic
    median_total_precip_median_lin + median_lag_2_prec_median_imp,
  data    = df,
  index   = c("id", "yday_corr"),      # unit.variable y time.variable
  force   = "two-way",                 # fijamos efectos fijos unitarios y temporales
  CV      = TRUE,                      # selección automática de r con cross‑validation
  r       = 1:4,                       # ≈ K = 4 en tu modelo original
  se      = TRUE,                      # errores estándar por bootstrap
  nboots  = 1000,                      # ajusta según tu tolerancia de tiempo/precisión
  parallel= F,                       # usa varios hilos si tu máquina lo permite
  method= "mc"
)
# ATT:
#                             ATT   S.E. CI.lower CI.upper p.value
# Tr obs equally weighted   0.223 0.1645 -0.09942   0.5455  0.1752
# Tr units equally weighted 0.223 0.1645 -0.09942   0.5455  0.1752
road.main.gsynth

plot(road.main.gsynth, main="-3 pre to +0, Quebec+Sherbrooke")

df2 <- subset(
  collisions_weather_corr_rect_quebec_synth,
  yday_corr <= (max_time - 7)            # mismo recorte temporal que en countSynth
)
df2$treat_unit <- df2$tr_contrl_corr
df2$post_treat <- ifelse(df2$yday_corr >= 35, 1, 0)
df2$D          <- df2$tr_contrl_corr * df2$post_treat   # interacción exposición x tiempo


road.main.gsynth2 <- fect::fect(
  formula = rate_lead2_lic ~ D + mean_min_temp_mean_lin + mean_max_temp_mean_lin +
    median_total_precip_median_lin + median_lag_2_prec_median_imp,
  data    = df2,
  index   = c("id", "yday_corr"),      # unit.variable y time.variable
  force   = "two-way",                 # fijamos efectos fijos unitarios y temporales
  CV      = TRUE,                      # selección automática de r con cross‑validation
  r       = 1:4,                       # ≈ K = 4 en tu modelo original
  se      = TRUE,                      # errores estándar por bootstrap
  nboots  = 1000,                      # ajusta según tu tolerancia de tiempo/precisión
  parallel= F,                       # usa varios hilos si tu máquina lo permite
  method= "mc"
)
cat("Quebec only as controls\n")
road.main.gsynth2
# ATT:
#                              ATT   S.E. CI.lower CI.upper p.value
# Tr obs equally weighted   0.2647 0.1803 -0.08866   0.6181   0.142
# Tr units equally weighted 0.2647 0.1803 -0.08866   0.6181   0.142

plot(road.main.gsynth2, main="-3 pre to +0, Quebec only")
```



### Forecast with Time GPT


We defined databases for `(t_int_man)` = `r (t_int_man)` days (df35) as pre-intervention period and a more restrictive `(t_int_man-3)` = `r (t_int_man-3)` (df33).

```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: timegpt
#| eval: true
#| result: "hold"
#| fig.cap: "Prediciton on the treated (adjusting on min max temperature and lagged precipitiation (2 days) and lagged rates"
#| fig-width: 9.5
#| fig-height: 5.5
#| fig-dpi: 600
#| results: "hold"
#https://cran.r-project.org/web/packages/nixtlar/vignettes/prediction-intervals.html
#https://cran.r-project.org/web/packages/nixtlar/vignettes/exogenous-variables.html
#https://cran.r-project.org/web/packages/nixtlar/vignettes/cross-validation.html
#https://cran.r-project.org/web/packages/nixtlar/vignettes/anomaly-detection.html

collisions_weather_corr_rect_synth_tr$Y_lag10 <- 
collisions_weather_corr_rect_synth_tr|> 
    mutate(Y_lag10 = dplyr::lag(rate_veh, 10, order_by = yday_corr))|>
    mutate(Y_lag10 = dplyr::lag(rate_veh, 10, order_by = yday_corr))|> 
    ungroup()|>
    pull(Y_lag10) 

collisions_weather_corr_rect_synth_tr$Y_lag10 <- 
collisions_weather_corr_rect_synth_tr|>
    mutate(
        third_value = nth(na.omit(Y_lag10), 1),
        Y_lag10 = ifelse(is.na(Y_lag10), third_value, Y_lag10)
    )|>
    dplyr::select(Y_lag10)|>
    ungroup()|>
    pull(Y_lag10)


df33 <- collisions_weather_corr_rect_synth_tr|> 
  arrange(year.x, yday_corr)|>                # 1,2,3… days relative to race
  filter(yday_corr<(t_int_man-3)) |> 
  transmute(
    unique_id = "treated",             # exactly this name
    ds        = as.POSIXct(as.Date("2019-05-19") + (yday_corr - 1)),
    y         = rate_veh,
    mean_median_lag_2_prec_median_imp         = mean_median_lag_2_prec_median_imp,
    mean_median_total_precip_median_lin= mean_median_total_precip_median_lin,
    mean_min_temp_mean_lin=  mean_min_temp_mean_lin,
    mean_max_temp_mean_lin= mean_max_temp_mean_lin,
    rate_lead1_lic= rate_lead1_lic,
    rate_lead2_lic= rate_lead2_lic,   
    sum_velocidad= sum_velocidad,
    sum_velocidad_lead1= sum_velocidad_lead1,
    sum_velocidad_lead2= sum_velocidad_lead2,
    exp_lic= exp_lic,    
    Y_lag10= Y_lag10
  )


df33c <- collisions_weather_corr_rect_synth_cntr|> 
  arrange(year.x, yday_corr)|>                # 1,2,3… days relative to race
  filter(yday_corr<(t_int_man-3)) |> 
  transmute(
    unique_id = "control",             # exactly this name
    ds        = as.POSIXct(as.Date("2019-05-19") + (yday_corr - 1)),
    y         = rate_veh,
    mean_median_lag_2_prec_median_imp         = mean_median_lag_2_prec_median_imp,
    mean_median_total_precip_median_lin= mean_median_total_precip_median_lin,
    mean_min_temp_mean_lin=  mean_min_temp_mean_lin,
    mean_max_temp_mean_lin= mean_max_temp_mean_lin,
    rate_lead1_lic= rate_lead1_lic,
    rate_lead2_lic= rate_lead2_lic,   
    sum_velocidad= sum_velocidad,
    sum_velocidad_lead1= sum_velocidad_lead1,
    sum_velocidad_lead2= sum_velocidad_lead2,
    exp_lic= exp_lic
  )

df33_post <- collisions_weather_corr_rect_synth_tr|> 
  arrange(year.x, yday_corr)|>                # 1,2,3… days relative to race
  filter(yday_corr>=(t_int_man-3), yday_corr<=max_time) |>                     # take the first 30
  transmute(
    unique_id = "treated",             # exactly this name
    ds        = as.POSIXct(as.Date("2019-05-19") + (yday_corr - 1)),
    y         = rate_veh,
    mean_median_lag_2_prec_median_imp         = mean_median_lag_2_prec_median_imp,
    mean_median_total_precip_median_lin= mean_median_total_precip_median_lin,
    mean_min_temp_mean_lin=  mean_min_temp_mean_lin,
    mean_max_temp_mean_lin= mean_max_temp_mean_lin,
    rate_lead1_lic= rate_lead1_lic,
    rate_lead2_lic= rate_lead2_lic,
    sum_velocidad= sum_velocidad,
    sum_velocidad_lead1= sum_velocidad_lead1,
    sum_velocidad_lead2= sum_velocidad_lead2,
    exp_lic= exp_lic,    
    Y_lag10= Y_lag10
  )

df33c_post <- collisions_weather_corr_rect_synth_cntr|> 
  arrange(year.x, yday_corr)|>                # 1,2,3… days relative to race
  filter(yday_corr>=(t_int_man-3), yday_corr<=max_time) |>                     # take the first 30
  transmute(
    unique_id = "control",             # exactly this name
    ds        = as.POSIXct(as.Date("2019-05-19") + (yday_corr - 1)),
    y         = rate_veh,
    mean_median_lag_2_prec_median_imp         = mean_median_lag_2_prec_median_imp,
    mean_median_total_precip_median_lin= mean_median_total_precip_median_lin,
    mean_min_temp_mean_lin=  mean_min_temp_mean_lin,
    mean_max_temp_mean_lin= mean_max_temp_mean_lin,
    rate_lead1_lic= rate_lead1_lic,
    rate_lead2_lic= rate_lead2_lic,
    sum_velocidad= sum_velocidad,
    sum_velocidad_lead1= sum_velocidad_lead1,
    sum_velocidad_lead2= sum_velocidad_lead2,
    exp_lic= exp_lic
  )

fc33 <- nixtla_client_forecast(
  rbind(mutate(df33[1:1,], ds=ds-1), df33),
  h     = 7,
  level = c(95),
  freq  = "D",          # <— this fixes the infer_frequency error
  finetune_steps = 10, #Number of steps used to finetune 'TimeGPT' in the new data.
  finetune_loss ="rmse"
)


p33 <- nixtla_client_plot(rbind(mutate(df33[1:1,], ds=ds-1), df33, df33_post), fc33,
                        max_insample_length = 100)


p33$layers[[1]]$aes_params$fill  <- "#eb8f8f"  # verde semitransparente
p33$layers[[1]]$aes_params$alpha <- 0.5


# 1) Banda de CI (layer 1)
p33$layers[[1]]$geom_params$fill  <- "#FFC10740"  # amarillo semitransparente
p33$layers[[1]]$geom_params$alpha <- 0.4

# 2) Línea histórica y pronóstico (layer 2)
# p$layers[[2]]$aes_params$colour  <- "steelblue"
p33$layers[[2]]$aes_params$linewidth <- 1.0


p33$layers[[1]]$aes_params$fill  <- "#eb8f8f"  # verde semitransparente
p33$layers[[1]]$aes_params$alpha <- 0.5

t0_33 <- min(p33$data$ds)
t0_33_posix <- min(p33$data$ds)
xint35   <- t0_33_posix + days((t_int_man-3))
t0_33 <- as.Date(min(p33$data$ds))

p33+
    scale_color_manual(
        values = c(
            # y        = "#FFD600",   # amarillo intenso
            TimeGPT  = "red"    # violeta (DarkOrchid / BlueViolet)
        ), 
        labels= c( TimeGPT = "Prediction (95% IC in gray area)" )
    ) +
    scale_x_datetime(
        name   = "Days of follow-up (33 days before race,\n7 days after the race)",
        limits = c(min(p33$data$ds), max(p33$data$ds)), 
        breaks = seq(min(p33$data$ds), max(p33$data$ds), by = "7 days"),
        labels = function(x) {
            # convierto POSIXct a Date y resto t0
            as.integer(as.Date(x) - t0_33 + 1)
        }
    ) +
    theme_classic(base_size = 13) +
    theme(
        strip.text   = element_blank(),   # quita el label de faceta
        strip.background = element_blank(),
        legend.position= "bottom"
    )+
    geom_vline(
        xintercept = xint35-days(2),
        linetype   = "dashed",
        color      = "black",
        linewidth       = 1
    ) +
    labs(y="Highspeed collisions per 1M license holders")
```

#### Only Quebec as controls


```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: timegpt_qc33
#| eval: true
#| fig.cap: "Prediction on the treated (-2 days before race, 7 days after — Québec)"
#| fig-width: 9.5
#| fig-height: 5.5
#| fig-dpi: 600
#| results: "hold"

## ── Conjuntos 33 días pre-intervención y 7 post ───────────────────────────────
df33_qc <- collisions_weather_corr_rect_quebec_synth_tr|> 
  arrange(year.x, yday_corr)|> 
  filter(yday_corr < (t_int_man-3))|> 
  transmute(
    unique_id = "treated",
    ds        = as.POSIXct(as.Date("2019-05-19") + (yday_corr - 1)),
    y         = rate_veh,
    mean_median_lag_2_prec_median_imp,
    mean_median_total_precip_median_lin,
    mean_min_temp_mean_lin,
    mean_max_temp_mean_lin,
    rate_lead1_lic,
    rate_lead2_lic,
    sum_velocidad= sum_velocidad,
    sum_velocidad_lead1= sum_velocidad_lead1,
    sum_velocidad_lead2= sum_velocidad_lead2,
    exp_lic= exp_lic
  )
df33c_qc <- collisions_weather_corr_rect_quebec_synth_cntr|> 
  arrange(year.x, yday_corr)|> 
  filter(yday_corr < (t_int_man-3))|> 
  transmute(
    unique_id = "control",
    ds        = as.POSIXct(as.Date("2019-05-19") + (yday_corr - 1)),
    y         = rate_veh,
    mean_median_lag_2_prec_median_imp,
    mean_median_total_precip_median_lin,
    mean_min_temp_mean_lin,
    mean_max_temp_mean_lin,
    rate_lead1_lic,
    rate_lead2_lic,
    sum_velocidad= sum_velocidad,
    sum_velocidad_lead1= sum_velocidad_lead1,
    sum_velocidad_lead2= sum_velocidad_lead2,
    exp_lic= exp_lic
  )

df33_post_qc <- collisions_weather_corr_rect_quebec_synth_tr|> 
  arrange(year.x, yday_corr)|> 
  filter(yday_corr >= (t_int_man-3), yday_corr <= max_time)|> 
  transmute(
    unique_id = "treated",
    ds        = as.POSIXct(as.Date("2019-05-19") + (yday_corr - 1)),
    y         = rate_veh,
    mean_median_lag_2_prec_median_imp,
    mean_median_total_precip_median_lin,
    mean_min_temp_mean_lin,
    mean_max_temp_mean_lin,
    rate_lead1_lic,
    rate_lead2_lic,
    sum_velocidad= sum_velocidad,
    sum_velocidad_lead1= sum_velocidad_lead1,
    sum_velocidad_lead2= sum_velocidad_lead2,
    exp_lic= exp_lic
  )

df33c_post_qc <- collisions_weather_corr_rect_quebec_synth_cntr|> 
  arrange(year.x, yday_corr)|> 
  filter(yday_corr >= (t_int_man-3), yday_corr <= max_time)|> 
  transmute(
    unique_id = "control",
    ds        = as.POSIXct(as.Date("2019-05-19") + (yday_corr - 1)),
    y         = rate_veh,
    mean_median_lag_2_prec_median_imp,
    mean_median_total_precip_median_lin,
    mean_min_temp_mean_lin,
    mean_max_temp_mean_lin,
    rate_lead1_lic,
    rate_lead2_lic,
    sum_velocidad= sum_velocidad,
    sum_velocidad_lead1= sum_velocidad_lead1,
    sum_velocidad_lead2= sum_velocidad_lead2,
    exp_lic= exp_lic
  )

## ── TimeGPT ───────────────────────────────────────────────────────────────────
fc33_qc <- nixtla_client_forecast(
  rbind(mutate(df33_qc[1:1,], ds = ds - 1), df33_qc),
  h     = 7,
  level = c(95),
  freq  = "D",
  finetune_steps = 10,
  finetune_loss  = "rmse"
)

p33_qc <- nixtla_client_plot(
  rbind(mutate(df33_qc[1:1,], ds = ds - 1),
        df33_qc,
        df33_post_qc),
  fc33_qc,
  max_insample_length = 100
)

## ── Estética idéntica a tu original ───────────────────────────────────────────
p33_qc$layers[[1]]$geom_params$fill   <- "#FFC10740"
p33_qc$layers[[1]]$geom_params$alpha  <- 0.4
p33_qc$layers[[2]]$aes_params$linewidth <- 1.0

t0_33_qc  <- as.Date(min(p33_qc$data$ds))
xint33_qc <- min(p33_qc$data$ds) + lubridate::days((t_int_man-3))

p33_qc +
  scale_color_manual(
    values = c(TimeGPT = "red"),
    labels = c(TimeGPT = "Prediction (95% IC in gray area)")
  ) +
  scale_x_datetime(
    name   = "Days of follow-up (33 before, 7 after)",
    limits = c(min(p33_qc$data$ds), max(p33_qc$data$ds)),
    breaks = seq(min(p33_qc$data$ds), max(p33_qc$data$ds), by = "7 days"),
    labels = function(x) as.integer(as.Date(x) - t0_33_qc + 1)
  ) +
  theme_classic(base_size = 13) +
  theme(
    strip.text       = element_blank(),
    strip.background = element_blank(),
    legend.position  = "bottom"
  ) +
  geom_vline(
    xintercept = xint33_qc,
    linetype   = "dashed",
    color      = "black",
    linewidth  = 1
  ) +
  labs(y = "High-speed collisions per 1 M license holders")
```


## Causal impact

We used the CausalImpact package, useful for time series data, inferring a counterfactual post race period based on pre-intervention components  [@Brodersen2015]. Given that the offset was constant, we used the sum of the high-speed collisions instead of rates.

### Quebec + Sherbrooke

#### -3 days pre

```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: causalimpact4
#| eval: true
#| fig.cap: "Causal impact (Local linear trend, -3 days race)"
#| fig-width: 9.5
#| fig-height: 5.5
#| fig-dpi: 600
#| results: "hold"

library(bsts)


# Add external regressors explicitly (e.g., temperatures, precipitation)
covariates33 <- cbind(
 # y = as.numeric(df35$y),
  mean_min_temp_mean_lin = df33$mean_min_temp_mean_lin,
  mean_max_temp_mean_lin = df33$mean_max_temp_mean_lin,
  mean_median_lag_2_prec_median_imp = df33$mean_median_lag_2_prec_median_imp,
  mean_median_total_precip_median_lin= df33$mean_median_total_precip_median_lin#,
  #Y_lag10 = df33$Y_lag10
)
# Ensure names explicitly
colnames(covariates33) <- c(#"y",
                          "mean_min_temp_mean_lin",
                          "mean_max_temp_mean_lin",
                          "mean_median_lag_2_prec_median_imp",
                          "mean_median_total_precip_median_lin"#,#"Y_lag10")
                          )
covariates33<- as.data.frame(covariates33)

covariates_future33 <- as.matrix(df33_post[1:7, c("mean_min_temp_mean_lin",
                                             "mean_max_temp_mean_lin",
                                             "mean_median_lag_2_prec_median_imp",
                                             "mean_median_total_precip_median_lin"#,
                                             #"Y_lag10"
                                             )])

covariates_full33 <- rbind(covariates33, covariates_future33)

lambda_opt <- forecast::BoxCox.lambda(as.numeric(df33$sum_velocidad_lead2), lower = -2, upper = 2)
y_bc <- BoxCox(as.numeric(df33$sum_velocidad_lead2), lambda_opt)

y33b <- c(as.numeric(df33$sum_velocidad_lead2), rep(NA, 7)) # respuesta escalada por 100 millones

y33b <- log(y33b)

y33b <- c(y_bc, rep(NA, 7)) # respuesta escalada por 100 millones

ss2b <- list()
ss2b <- AddLocalLinearTrend(ss2b, y33b)
ss2b <- AddSeasonal(ss2b, y33b, nseasons = 7, season.duration = 1)
#ss2b <- AddAutoAr(ss2b, y33b, lags = 1)


ss2b <- AddDynamicRegression(ss2b, covariates_full33)

set.seed(2125)
model33b <- bsts(y33b, state.specification = ss2b, seed=2125,
              niter = 4e4, family = "gaussian")

#plot(model, "components", burn = 2000)

impact3d_ratio_resp2 <- CausalImpact(bsts.model = model33b,model.args = list(prior.level.sd=.1, dynamic.regression=T),
              post.period.response = as.numeric(BoxCox(as.numeric(df33_post$sum_velocidad_lead2[1:7]), lambda_opt)))
                #log(df33_post$sum_velocidad_lead2[1:7]))
#plot(impact3d, "original") 

summary(impact3d_ratio_resp2)

plot(impact3d_ratio_resp2)



err_ci_00 <- impact3d_ratio_resp2$series$response - impact3d_ratio_resp2$series$point.pred
rmse_ci_00 <- sqrt(mean(err_ci_00[1:35]^2))
mean_pre_00 <- mean(df33$sum_velocidad_lead2[1:35]) 
sd_pre_00 <- sd(df33$sum_velocidad_lead2[1:35])
nrmse_00 <- rmse_ci_00 / mean_pre_00 # RMSE relativo; 
sd_ratio_00 <- rmse_ci_00 / sd_pre_00 # RMSE relativo a la variabilidad natural

cat(sprintf("RMSE on 35 post-period points: %.2f (average absolute forecast error)\n", rmse_ci_00))
cat(sprintf("NRMSE: %.2f (RMSE relative to pre-period mean, %.1f)\n", nrmse_00, mean_pre_00))
cat(sprintf("SD-ratio: %.2f (RMSE vs pre-period variability, SD=%.1f)\n", sd_ratio_00, sd_pre_00))

# Benchmark naïve (random walk de un paso):
naive_rmse_00 <- sqrt(mean((df33$sum_velocidad_lead2[2:35] - df33$sum_velocidad_lead2[1:34])^2))
improvement_00 <- 1 - rmse_ci_00 / naive_rmse_00  # >0 = mejora sobre el naïve

```



```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: causalimpact4-c
#| eval: true
#| fig.cap: "Causal impact (Local linear trend, -3 days race), on Controls (Sherbrook+Quebec)"
#| fig-width: 9.5
#| fig-height: 5.5
#| fig-dpi: 600
#| results: "hold"

# Add external regressors explicitly (e.g., temperatures, precipitation)
covariates33c <- cbind(
 # y = as.numeric(df35$y),
  mean_min_temp_mean_lin = df33c$mean_min_temp_mean_lin,
  mean_max_temp_mean_lin = df33c$mean_max_temp_mean_lin,
  mean_median_lag_2_prec_median_imp = df33c$mean_median_lag_2_prec_median_imp,
  mean_median_total_precip_median_lin= df33c$mean_median_total_precip_median_lin#,
  #Y_lag10 = df33$Y_lag10
)
# Ensure names explicitly
colnames(covariates33c) <- c(#"y",
                          "mean_min_temp_mean_lin",
                          "mean_max_temp_mean_lin",
                          "mean_median_lag_2_prec_median_imp",
                          "mean_median_total_precip_median_lin"#,#"Y_lag10")
                          )
covariates33c<- as.data.frame(covariates33c)

covariates_future33c <- as.matrix(df33c_post[1:7, c("mean_min_temp_mean_lin",
                                             "mean_max_temp_mean_lin",
                                             "mean_median_lag_2_prec_median_imp",
                                             "mean_median_total_precip_median_lin"#,
                                             #"Y_lag10"
                                             )])

covariates_full33c <- rbind(covariates33c, covariates_future33c)

# y33cb <- c(as.numeric(df33c$sum_velocidad_lead2), rep(NA, 7)) # respuesta escalada por 100 millones
# y33cb <- log(y33cb)

lambda_opt <- forecast::BoxCox.lambda(as.numeric(df33c$sum_velocidad_lead2), lower = -2, upper = 2)
y_bc <- BoxCox(as.numeric(df33c$sum_velocidad_lead2), lambda_opt)
y33cb <- c(y_bc, rep(NA, 7)) # respuesta escalada por 100 millones


ss2cb <- list()
ss2cb <- AddLocalLinearTrend(ss2cb, y33cb)
ss2cb <- AddSeasonal(ss2cb, y33cb, nseasons = 7, season.duration = 1)
#ss2b <- AddAutoAr(ss2b, y33cb, lags = 1)


ss2cb <- AddDynamicRegression(ss2cb, covariates_full33c)

set.seed(2125)
model33cb <- bsts(y33cb, state.specification = ss2cb, seed=2125,
              niter = 4e4, family = "gaussian")

#plot(model, "components", burn = 2000)

impact3d_ratio_resp2_c <- CausalImpact(bsts.model = model33cb,model.args = list(prior.level.sd=.1, dynamic.regression=T),
              post.period.response = as.numeric(BoxCox(as.numeric(df33c_post$sum_velocidad_lead2[1:7]), lambda_opt)))
                #log(df33c_post$sum_velocidad_lead2[1:7]))
#plot(impact3d, "original") 

summary(impact3d_ratio_resp2_c)

plot(impact3d_ratio_resp2_c)




err_ci_0c <- impact3d_ratio_resp2_c$series$response - impact3d_ratio_resp2_c$series$point.pred
rmse_ci_0c <- sqrt(mean(err_ci_0c[1:35]^2))
mean_pre_0c <- mean(df33c$sum_velocidad_lead2[1:35]) 
sd_pre_0c <- sd(df33c$sum_velocidad_lead2[1:35])
nrmse_0c <- rmse_ci_0c / mean_pre_0c # RMSE relativo; 
sd_ratio_0c <- rmse_ci_0c / sd_pre_0c # RMSE relativo a la variabilidad natural

cat(sprintf("RMSE on 35 post-period points: %.2f (average absolute forecast error)\n", rmse_ci_0c))
cat(sprintf("NRMSE: %.2f (RMSE relative to pre-period mean, %.1f)\n", nrmse_0c, mean_pre_0c))
cat(sprintf("SD-ratio: %.2f (RMSE vs pre-period variability, SD=%.1f)\n", sd_ratio_0c, sd_pre_0c))

# Benchmark naïve (random walk de un paso):
naive_rmse_0c <- sqrt(mean((df33c$sum_velocidad_lead2[2:35] - df33c$sum_velocidad_lead2[1:34])^2))
improvement_0c <- 1 - rmse_ci_0c / naive_rmse_0c  # >0 = mejora sobre el naïve

```


#### -3 days pre, +0

```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: causalimpact3b
#| eval: true
#| fig.cap: "Causal impact (Main, -3 days race, +0 days post)"
#| fig-width: 9.5
#| fig-height: 5.5
#| fig-dpi: 600
#| results: "hold"

# yss2ab<- c(as.numeric(df33$sum_velocidad_lead2), rep(NA,3))
# yss2ab<- log(yss2ab)
lambda_opt <- forecast::BoxCox.lambda(as.numeric(df33$sum_velocidad_lead2), lower = -2, upper = 2)
y_bc <- BoxCox(as.numeric(df33$sum_velocidad_lead2), lambda_opt)
yss2ab <- c(y_bc, rep(NA, 3)) # respuesta escalada por 100 millones

# Model 2
ss2ab <- list()
# Local trend, weekly-seasonal #https://qastack.mx/stats/209426/predictions-from-bsts-model-in-r-are-failing-completely - PUSE UN GENERALIZED LOCAL TREND
ss2ab <- AddLocalLevel(ss2ab, yss2ab) #
# Add weekly seasonal
ss2ab <- AddSeasonal(ss2ab, yss2ab, 
                    nseasons=7, season.duration = 1) #weeks OJO, ESTOS NO SON WEEKS VERDADEROS. PORQUE TENGO MAS DE EUN AÑO
#ss2ab <- AddAutoAr(ss2ab, y = c(as.numeric(df33$rate_lead2_lic), rep(NA,3)), lags = 10) #NO PUEDO AREGAR AR1 CON gaussian
# For example, to add a day-of-week component to data with daily granularity, use model.args = list(nseasons = 7, season.duration = 1). To add a day-of-week component to data with hourly granularity, set model.args = list(nseasons = 7, season.duration = 24).

covariates_future33b <- as.matrix(df33_post[1:3, c("mean_min_temp_mean_lin",
                                             "mean_max_temp_mean_lin",
                                             "mean_median_lag_2_prec_median_imp",
                                             "mean_median_total_precip_median_lin"#,
                                             #"Y_lag10"
                                             )])

covariates_full33b <- rbind(covariates33, covariates_future33b)

# Include regression terms
ss2ab <- AddDynamicRegression(ss2ab, covariates_full33b)

model2ab_ratio <- bsts(yss2ab,
               state.specification = ss2ab, #A list with elements created by AddLocalLinearTrend, AddSeasonal, and similar functions for adding components of state. See the help page for state.specification.
               family ="gaussian", #A Bayesian Analysis of Time-Series Event Count Data
               niter = 4e4, 
               #burn = 2000, #http://finzi.psych.upenn.edu/library/bsts/html/SuggestBurn.html Suggest the size of an MCMC burn in sample as a proportion of the total run.
               seed= 2125)

## Since we have predictor variables in the model, we need to explicitly
# make their coefficients time-varying using AddDynamicRegression().

impact3ab_ratio_resp <- CausalImpact(bsts.model = model2ab_ratio,model.args = list(prior.level.sd=.1, dynamic.regression=T),
                                                   post.period.response = as.numeric(BoxCox(as.numeric(df33_post$sum_velocidad_lead2[1:3]), lambda_opt)))
              #post.period.response = log(df33_post$sum_velocidad_lead2[1:3]))

summary(impact3ab_ratio_resp)
plot(impact3ab_ratio_resp)

err_ci <- impact3ab_ratio_resp$series$response - impact3ab_ratio_resp$series$point.pred
rmse_ci <- sqrt(mean(err_ci[1:34]^2))
mean_pre <- mean(df33$sum_velocidad_lead2[1:34]) 
sd_pre <- sd(df33$sum_velocidad_lead2[1:34])
nrmse <- rmse_ci / mean_pre # RMSE relativo; 
sd_ratio <- rmse_ci / sd_pre # RMSE relativo a la variabilidad natural

cat(sprintf("RMSE on 35 post-period points: %.2f (average absolute forecast error)\n", rmse_ci))
cat(sprintf("NRMSE: %.2f (RMSE relative to pre-period mean, %.1f)\n", nrmse, mean_pre))
cat(sprintf("SD-ratio: %.2f (RMSE vs pre-period variability, SD=%.1f)\n", sd_ratio, sd_pre))

# Benchmark naïve (random walk de un paso):
naive_rmse <- sqrt(mean((df33$sum_velocidad_lead2[2:34] - df33$sum_velocidad_lead2[1:33])^2))
improvement <- 1 - rmse_ci / naive_rmse  # >0 = mejora sobre el naïve

```



```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: causalimpact3b-c
#| eval: true
#| fig.cap: "Causal impact (Main, -3 days race, +0 days post)"
#| fig-width: 9.5
#| fig-height: 5.5
#| fig-dpi: 600
#| results: "hold"


# yss2cab <- c(as.numeric(df33c$sum_velocidad_lead2), rep(NA,3))
# yss2cab <- log(yss2cab)
lambda_opt <- forecast::BoxCox.lambda(as.numeric(df33c$sum_velocidad_lead2), lower = -2, upper = 2)
y_bc <- BoxCox(as.numeric(df33c$sum_velocidad_lead2), lambda_opt)
yss2cab <- c(y_bc, rep(NA, 3)) # respuesta escalada por 100 millones

# Model 2
ss2cab <- list()
# Local trend, weekly-seasonal #https://qastack.mx/stats/209426/predictions-from-bsts-model-in-r-are-failing-completely - PUSE UN GENERALIZED LOCAL TREND
ss2cab <- AddLocalLevel(ss2cab, yss2cab) #
# Add weekly seasonal
ss2cab <- AddSeasonal(ss2cab, yss2cab, 
                    nseasons=7, season.duration = 1) #weeks OJO, ESTOS NO SON WEEKS VERDADEROS. PORQUE TENGO MAS DE EUN AÑO
#ss2ab <- AddAutoAr(ss2ab, y = c(as.numeric(df33$rate_lead2_lic), rep(NA,3)), lags = 10) #NO PUEDO AREGAR AR1 CON gaussian
# For example, to add a day-of-week component to data with daily granularity, use model.args = list(nseasons = 7, season.duration = 1). To add a day-of-week component to data with hourly granularity, set model.args = list(nseasons = 7, season.duration = 24).

covariates_future33cb <- as.matrix(df33c_post[1:3, c("mean_min_temp_mean_lin",
                                             "mean_max_temp_mean_lin",
                                             "mean_median_lag_2_prec_median_imp",
                                             "mean_median_total_precip_median_lin"#,
                                             #"Y_lag10"
                                             )])

covariates_full33cb <- rbind(covariates33c, covariates_future33cb)

# Include regression terms
ss2cab <- AddDynamicRegression(ss2cab, covariates_full33cb)

model2cab_ratio <- bsts(yss2cab,
               state.specification = ss2cab, #A list with elements created by AddLocalLinearTrend, AddSeasonal, and similar functions for adding components of state. See the help page for state.specification.
               family ="gaussian", #A Bayesian Analysis of Time-Series Event Count Data
               niter = 4e4, 
               #burn = 2000, #http://finzi.psych.upenn.edu/library/bsts/html/SuggestBurn.html Suggest the size of an MCMC burn in sample as a proportion of the total run.
               seed= 2125)

## Since we have predictor variables in the model, we need to explicitly
# make their coefficients time-varying using AddDynamicRegression().

impact3cab_ratio_resp <- CausalImpact(bsts.model = model2cab_ratio,model.args = list(prior.level.sd=.1, dynamic.regression=T),
              #post.period.response = log(df33c_post$sum_velocidad_lead2[1:3]))
              post.period.response = as.numeric(BoxCox(as.numeric(df33c_post$sum_velocidad_lead2[1:3]), lambda_opt)))

summary(impact3cab_ratio_resp)
plot(impact3cab_ratio_resp)

err_ci_c <- impact3cab_ratio_resp$series$response - impact3cab_ratio_resp$series$point.pred
rmse_ci_c <- sqrt(mean(err_ci_c[1:34]^2))
mean_pre_c <- mean(df33c$sum_velocidad_lead2[1:34]) 
sd_pre_c <- sd(df33c$sum_velocidad_lead2[1:34])
nrmse_c <- rmse_ci_c / mean_pre_c # RMSE relativo; 
sd_ratio_c <- rmse_ci_c / sd_pre_c # RMSE relativo a la variabilidad natural

cat(sprintf("RMSE on 35 post-period points: %.2f (average absolute forecast error)\n", rmse_ci_c))
cat(sprintf("NRMSE: %.2f (RMSE relative to pre-period mean, %.1f)\n", nrmse_c, mean_pre_c))
cat(sprintf("SD-ratio: %.2f (RMSE vs pre-period variability, SD=%.1f)\n", sd_ratio_c, sd_pre_c))

# Benchmark naïve (random walk de un paso):
naive_rmse_c <- sqrt(mean((df33c$sum_velocidad_lead2[2:34] - df33c$sum_velocidad_lead2[1:33])^2))
improvement_c <- 1 - rmse_ci_c / naive_rmse_c  # >0 = mejora sobre el naïve

```

```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: causalimpact4b
#| eval: true
#| fig.cap: "Causal impact (Local linear trend, -3 days race, 0)"
#| fig-width: 9.5
#| fig-height: 5.5
#| fig-dpi: 600
#| results: "hold"

# y33b <- c(as.numeric(df33$sum_velocidad_lead2), rep(NA, 3)) # respuesta escalada por 100 millones
# y33b <- log(y33b)
lambda_opt <- forecast::BoxCox.lambda(as.numeric(df33$sum_velocidad_lead2), lower = -2, upper = 2)
y_bc <- BoxCox(as.numeric(df33$sum_velocidad_lead2), lambda_opt)
y33b <- c(y_bc, rep(NA, 3)) # respuesta escalada por 100 millones


ss2b <- list()
ss2b <- AddLocalLinearTrend(ss2b, y33b)
ss2b <- AddSeasonal(ss2b, y33b, nseasons = 7, season.duration = 1)
#ss2b <- AddAutoAr(ss2b, y33b, lags = 1)


ss2b <- AddDynamicRegression(ss2b, covariates_full33b)

set.seed(2125)
model33b <- bsts(y33b, state.specification = ss2b, seed=2125,
              niter = 4e4, family = "gaussian")

#plot(model, "components", burn = 2000)

impact3db_ratio_resp2 <- CausalImpact(bsts.model = model33b, model.args = list(prior.level.sd=.1, dynamic.regression=T),
              #post.period.response = log(df33_post$sum_velocidad_lead2[1:3]))
              post.period.response = as.numeric(BoxCox(as.numeric(df33_post$sum_velocidad_lead2[1:3]), lambda_opt)))

summary(impact3db_ratio_resp2)

plot(impact3db_ratio_resp2)


err_ci_corr <- impact3db_ratio_resp2$series$response - impact3db_ratio_resp2$series$point.pred
rmse_ci_corr <- sqrt(mean(err_ci_corr[1:34]^2))
mean_pre_corr <- mean(df33$sum_velocidad_lead2[1:34]) 
sd_pre_corr <- sd(df33$sum_velocidad_lead2[1:34])
nrmse_corr <- rmse_ci_corr / mean_pre_corr # RMSE relativo; 
sd_ratio_corr <- rmse_ci_corr / sd_pre_corr # RMSE relativo a la variabilidad natural

cat(sprintf("RMSE on 35 post-period points: %.2f (average absolute forecast error)\n", rmse_ci_corr))
cat(sprintf("NRMSE: %.2f (RMSE relative to pre-period mean, %.1f)\n", nrmse_corr, mean_pre_corr))
cat(sprintf("SD-ratio: %.2f (RMSE vs pre-period variability, SD=%.1f)\n", sd_ratio_corr, sd_pre_corr))
```



```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: causalimpact4b-c
#| eval: true
#| fig.cap: "Causal impact (Local linear trend, -3 days race, 0), controls"
#| fig-width: 9.5
#| fig-height: 5.5
#| fig-dpi: 600
#| results: "hold"

# y33b_c <- c(as.numeric(df33c$sum_velocidad_lead2), rep(NA, 3)) # respuesta escalada por 100 millones
# y33b_c <- log(y33b_c)
lambda_opt <- forecast::BoxCox.lambda(as.numeric(df33c$sum_velocidad_lead2), lower = -2, upper = 2)
y_bc <- BoxCox(as.numeric(df33c$sum_velocidad_lead2), lambda_opt)
y33b_c <- c(y_bc, rep(NA, 3)) # respuesta escalada por 100 millones


ss2b_c <- list()
ss2b_c <- AddLocalLinearTrend(ss2b_c, y33b_c)
ss2b_c <- AddSeasonal(ss2b_c, y33b_c, nseasons = 7, season.duration = 1)
#ss2b <- AddAutoAr(ss2b, y33b, lags = 1)


covariates_future33c2 <- as.matrix(df33c_post[1:3, c("mean_min_temp_mean_lin",
                                             "mean_max_temp_mean_lin",
                                             "mean_median_lag_2_prec_median_imp",
                                             "mean_median_total_precip_median_lin"#,
                                             #"Y_lag10"
                                             )])

covariates_full33c2 <- rbind(covariates33c, covariates_future33c2)

ss2b_c <- AddDynamicRegression(ss2b_c, covariates_full33c2)

set.seed(2125)
model33b_c <- bsts(y33b_c, state.specification = ss2b_c, seed=2125,
              niter = 4e4, family = "gaussian")

#plot(model, "components", burn = 2000)

impact3db_ratio_resp2_c <- CausalImpact(bsts.model = model33b_c, model.args = list(prior.level.sd=.1, dynamic.regression=T),
              #post.period.response = log(df33c_post$sum_velocidad_lead2[1:3]))
              post.period.response = as.numeric(BoxCox(as.numeric(df33c_post$sum_velocidad_lead2[1:3]), lambda_opt)))

summary(impact3db_ratio_resp2_c)

plot(impact3db_ratio_resp2_c)


err_ci_c_corr <- impact3db_ratio_resp2_c$series$response - impact3db_ratio_resp2_c$series$point.pred
rmse_ci_c_corr <- sqrt(mean(err_ci_c_corr[1:34]^2))
mean_pre_c_corr <- mean(df33c$sum_velocidad_lead2[1:34]) 
sd_pre_c_corr <- sd(df33c$sum_velocidad_lead2[1:34])
nrmse_c_corr <- rmse_ci_c_corr / mean_pre_c_corr # RMSE relativo; 
sd_ratio_c_corr <- rmse_ci_c_corr / sd_pre_c_corr # RMSE relativo a la variabilidad natural

cat(sprintf("RMSE on 35 post-period points: %.2f (average absolute forecast error)\n", rmse_ci_c_corr))
cat(sprintf("NRMSE: %.2f (RMSE relative to pre-period mean, %.1f)\n", nrmse_c_corr, mean_pre_c_corr))
cat(sprintf("SD-ratio: %.2f (RMSE vs pre-period variability, SD=%.1f)\n", sd_ratio_c_corr, sd_pre_c_corr))
```


### Only Quebec as controls


#### -3 days pre

```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: causalimpact_qc33
#| eval: true
#| fig.cap: "Causal impact (Québec only -3 days pre, model w/ local level)"
#| fig-width: 9.5
#| fig-height: 5.5
#| fig-dpi: 600
#| results: "hold"
#| 
## ── Vector de respuesta ───────────────────────────────────────────────────────
# y33_qc <- c(as.numeric(df33_qc$sum_velocidad_lead2), rep(NA, 3))  # 7 post-días + 2 de ajuste
# y33_qc <- log(y33_qc)
lambda_opt <- forecast::BoxCox.lambda(as.numeric(df33_qc$sum_velocidad_lead2), lower = -2, upper = 2)
y_bc <- BoxCox(as.numeric(df33_qc$sum_velocidad_lead2), lambda_opt)
y33_qc <- c(y_bc, rep(NA, 3)) # respuesta escalada por 100 millones


## ── Covariables (data.frame) ─────────────────────────────────────────────────
covariates33_qc <- df33_qc |> dplyr::select(
  mean_min_temp_mean_lin,
  mean_max_temp_mean_lin,
  mean_median_lag_2_prec_median_imp,
  mean_median_total_precip_median_lin#,
  #Y_lag10
)

covariates_future33_qc <- df33_post_qc |> dplyr::select(
  mean_min_temp_mean_lin,
  mean_max_temp_mean_lin,
  mean_median_lag_2_prec_median_imp,
  mean_median_total_precip_median_lin#,
  #Y_lag10
)

covariates_full33_qc <- dplyr::bind_rows(covariates33_qc, covariates_future33_qc[1:7,])
covariates_full33_qc2 <- dplyr::bind_rows(covariates33_qc, covariates_future33_qc[1:3,])

```



```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: causalimpact_qc33-c
#| eval: true
#| fig.cap: "Causal impact (Québec only -3 days pre, model w/ local level)"
#| fig-width: 9.5
#| fig-height: 5.5
#| fig-dpi: 600
#| results: "hold"
#| 
## ── Vector de respuesta ───────────────────────────────────────────────────────
# y33_qc_c <- c(as.numeric(df33c_qc$sum_velocidad_lead2), rep(NA, 3))  # 7 post-días + 2 de ajuste
# y33_qc_c <- log(y33_qc_c)
lambda_opt <- forecast::BoxCox.lambda(as.numeric(df33c_qc$sum_velocidad_lead2), lower = -2, upper = 2)
y_bc <- BoxCox(as.numeric(df33c_qc$sum_velocidad_lead2), lambda_opt)
y33_qc_c <- c(y_bc, rep(NA, 3)) # respuesta escalada por 100 millones

## ── Covariables (data.frame) ─────────────────────────────────────────────────
covariates33_qc_c <- df33c_qc |> dplyr::select(
  mean_min_temp_mean_lin,
  mean_max_temp_mean_lin,
  mean_median_lag_2_prec_median_imp,
  mean_median_total_precip_median_lin#,
  #Y_lag10
)

covariates_future33_qc_c <- df33c_post_qc |> dplyr::select(
  mean_min_temp_mean_lin,
  mean_max_temp_mean_lin,
  mean_median_lag_2_prec_median_imp,
  mean_median_total_precip_median_lin#,
  #Y_lag10
)

covariates_full33_qc_c <- dplyr::bind_rows(covariates33_qc_c, covariates_future33_qc_c[1:7,])
covariates_full33_qc_c2 <- dplyr::bind_rows(covariates33_qc_c, covariates_future33_qc_c[1:3,])

```

#### -3 days pre, +0

```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: causalimpact_qc33b
#| eval: true
#| fig.cap: "Causal impact (Québec only -3 days pre, 0 days post, model w/ local level)"
#| fig-width: 9.5
#| fig-height: 5.5
#| fig-dpi: 600
#| results: "hold"

## ── Vector de respuesta ───────────────────────────────────────────────────────
# y33b_qc <- c(as.numeric(df33_qc$sum_velocidad_lead2), rep(NA, 3))  # 7 post-días + 2 de ajuste
# y33b_qc <- log(y33b_qc)
lambda_opt <- forecast::BoxCox.lambda(as.numeric(df33_qc$sum_velocidad_lead2), lower = -2, upper = 2)
y_bc <- forecast::BoxCox(as.numeric(df33_qc$sum_velocidad_lead2), lambda_opt)
y33b_qc <- c(y_bc, rep(NA, 3)) # respuesta escalada por 100 millones

covariates_full33b_qc <- dplyr::bind_rows(covariates33_qc, covariates_future33_qc[1:3,])

## ── Especificación del estado: nivel local ───────────────────────────────────
ss2ab_qc <- list()
ss2ab_qc <- AddLocalLevel(ss2ab_qc, y33b_qc)
ss2ab_qc <- AddSeasonal(ss2ab_qc, y33b_qc, nseasons = 7, season.duration = 1)
#ss2ab_qc <- AddAutoAr   (ss2ab_qc, y33b_qc, lags = 10)
ss2ab_qc <- AddDynamicRegression(ss2ab_qc, covariates_full33b_qc)

## ── Ajuste BSTS y CausalImpact ───────────────────────────────────────────────
model2ab_qc <- bsts(
  y33b_qc,
  state.specification = ss2ab_qc,
  family = "gaussian",
  niter  = 4e4,
  seed   = 2125
)

impact33b_qc <- CausalImpact(
  bsts.model           = model2ab_qc,
  model.args           = list(prior.level.sd = .1, dynamic.regression = TRUE),
  #post.period.response = log(df33_post_qc$sum_velocidad_lead2[1:3])
  post.period.response = as.numeric(forecast::BoxCox(as.numeric(df33_post_qc$sum_velocidad_lead2[1:3]), lambda_opt))
)

summary(impact33b_qc)


err_ci_qb_corr <- impact33b_qc$series$response - impact33b_qc$series$point.pred
rmse_ci_qb_corr <- sqrt(mean(err_ci_qb_corr[1:35]^2))
mean_pre_qb_corr <- mean(df33_qc$sum_velocidad_lead2[1:35], na.rm=T) 
sd_pre_qb_corr <- sd(df33_qc$sum_velocidad_lead2[1:35], na.rm=T)
nrmse_qb_corr <- rmse_ci_qb_corr / mean_pre_qb_corr # RMSE relativo; 
sd_ratio_qb_corr <- rmse_ci_qb_corr / sd_pre_qb_corr # RMSE relativo a la variabilidad natural

cat(sprintf("RMSE on 35 post-period points: %.2f (average absolute forecast error)\n", rmse_ci_qb_corr))
cat(sprintf("NRMSE: %.2f (RMSE relative to pre-period mean, %.1f)\n", nrmse_qb_corr, mean_pre_qb_corr))
cat(sprintf("SD-ratio: %.2f (RMSE vs pre-period variability, SD=%.1f)\n", sd_ratio_qb_corr, sd_pre_qb_corr))


plot(impact33b_qc)
```



```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: causalimpact_qc33b-c
#| eval: true
#| fig.cap: "Causal impact (Québec only -3 days pre, 0 days post, model w/ local level)"
#| fig-width: 9.5
#| fig-height: 5.5
#| fig-dpi: 600
#| results: "hold"
#| 
## ── Vector de respuesta ───────────────────────────────────────────────────────
# y33b_qc_c <- c(as.numeric(df33c_qc$sum_velocidad_lead2), rep(NA, 3))  # 7 post-días + 2 de ajuste
# y33b_qc_c <- log(y33b_qc_c)
lambda_opt <- forecast::BoxCox.lambda(as.numeric(df33c_qc$sum_velocidad_lead2), lower = -2, upper = 2)
y33b_qc_c <- BoxCox(as.numeric(df33c_qc$sum_velocidad_lead2), lambda_opt)
y33b_qc_c <- c(as.numeric(y33b_qc_c), rep(NA, 3)) 

## ── Especificación del estado: nivel local ───────────────────────────────────
ss2ab_qc_c <- list()
ss2ab_qc_c <- AddLocalLevel(ss2ab_qc_c, y33b_qc_c)
ss2ab_qc_c <- AddSeasonal(ss2ab_qc_c, y33b_qc_c, nseasons = 7, season.duration = 1)
#ss2ab_qc <- AddAutoAr   (ss2ab_qc, y33b_qc, lags = 10)
ss2ab_qc_c <- AddDynamicRegression(ss2ab_qc_c, covariates_full33_qc_c2)

## ── Ajuste BSTS y CausalImpact ───────────────────────────────────────────────
model2ab_qc_c <- bsts(
  y33b_qc_c,
  state.specification = ss2ab_qc_c,
  family = "gaussian",
  niter  = 4e4,
  seed   = 2125
)

impact33ab_qc_c <- CausalImpact(
  bsts.model           = model2ab_qc_c,
  model.args           = list(prior.level.sd = .1, dynamic.regression = TRUE),
  #post.period.response = log(df33c_post_qc$sum_velocidad_lead2[1:3])
  post.period.response = as.numeric(BoxCox(as.numeric(df33c_post_qc$sum_velocidad_lead2[1:3]), lambda_opt))
)


summary(impact33ab_qc_c)

err_ci_c_qb_corr <- impact33ab_qc_c$series$response - impact33ab_qc_c$series$point.pred
rmse_ci_c_qb_corr <- sqrt(mean(err_ci_c_qb_corr[1:34]^2))
mean_pre_c_qb_corr <- mean(df33c_qc$sum_velocidad_lead2[1:34], na.rm=T) 
sd_pre_c_qb_corr <- sd(df33c_qc$sum_velocidad_lead2[1:34], na.rm=T)
nrmse_c_qb_corr <- rmse_ci_c_qb_corr / mean_pre_c_qb_corr # RMSE relativo; 
sd_ratio_c_qb_corr <- rmse_ci_c_qb_corr / sd_pre_c_qb_corr # RMSE relativo a la variabilidad natural

cat(sprintf("RMSE on 35 post-period points: %.2f (average absolute forecast error)\n", rmse_ci_c_qb_corr))
cat(sprintf("NRMSE: %.2f (RMSE relative to pre-period mean, %.1f)\n", nrmse_c_qb_corr, mean_pre_c_qb_corr))
cat(sprintf("SD-ratio: %.2f (RMSE vs pre-period variability, SD=%.1f)\n", sd_ratio_c_qb_corr, sd_pre_c_qb_corr))

plot(impact33ab_qc_c)
```


```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: causalimpact2_qc33b
#| eval: true
#| fig.cap: "Causal impact (Québec only,-3 days pre, local linear trend)"
#| fig-width: 9.5
#| fig-height: 5.5
#| fig-dpi: 600
#| results: "hold"

## ── Vector de respuesta ───────────────────────────────────────────────────────
# y33b_lin_qc <- c(as.numeric(df33_qc$sum_velocidad_lead2), rep(NA, 3))
# y33b_lin_qc <- log(y33b_lin_qc)
lambda_opt <- forecast::BoxCox.lambda(as.numeric(df33_qc$sum_velocidad_lead2), lower = -2, upper = 2)
y33b_lin_qc <- BoxCox(as.numeric(df33_qc$sum_velocidad_lead2), lambda_opt)
y33b_lin_qc <- c(as.numeric(y33b_lin_qc), rep(NA, 3))

## ── Covariables (data.frame) ─────────────────────────────────────────────────
covariates_full33b_qc <- dplyr::bind_rows(covariates33_qc, covariates_future33_qc)

## ── Estado: tendencia local lineal ───────────────────────────────────────────
ss2bb_qc <- list()
ss2bb_qc <- AddLocalLinearTrend(ss2bb_qc, y33b_lin_qc)
ss2bb_qc <- AddSeasonal        (ss2bb_qc, y33b_lin_qc, nseasons = 7, season.duration = 1)
#ss2bb_qc <- AddAutoAr          (ss2bb_qc, y33b_lin_qc, lags = 1)
ss2bb_qc <- AddDynamicRegression(ss2bb_qc, covariates_full33b_qc)

## ── Ajuste BSTS y CausalImpact ───────────────────────────────────────────────
model33b_qc2 <- bsts(
  y33b_lin_qc,
  state.specification = ss2bb_qc,
  family = "gaussian",
  niter  = 4e4,
  seed   = 2125
)

impact2_33b_qc2 <- CausalImpact(
  bsts.model           = model33b_qc2,
  model.args           = list(prior.level.sd = .1, dynamic.regression = TRUE),
  #post.period.response = log(df33_post_qc$sum_velocidad_lead2[1:3]),
  post.period.response = as.numeric(forecast::BoxCox(as.numeric(df33_post_qc$sum_velocidad_lead2[1:3]), lambda_opt))
)

summary(impact2_33b_qc2)

err_ci_qb_corr2 <- impact2_33b_qc2$series$response - impact2_33b_qc2$series$point.pred
rmse_ci_qb_corr2 <- sqrt(mean(err_ci_qb_corr2[1:34]^2))
mean_pre_qb_corr2 <- mean(df33_qc$sum_velocidad_lead2[1:34], na.rm=T) 
sd_pre_qb_corr2 <- sd(df33_qc$sum_velocidad_lead2[1:34], na.rm=T)
nrmse_qb_corr2 <- rmse_ci_qb_corr2 / mean_pre_qb_corr2 # RMSE relativo; 
sd_ratio_qb_corr2 <- rmse_ci_qb_corr2 / sd_pre_qb_corr2 # RMSE relativo a la variabilidad natural

cat(sprintf("RMSE on 35 post-period points: %.2f (average absolute forecast error)\n", rmse_ci_qb_corr2))
cat(sprintf("NRMSE: %.2f (RMSE relative to pre-period mean, %.1f)\n", nrmse_qb_corr2, mean_pre_qb_corr2))
cat(sprintf("SD-ratio: %.2f (RMSE vs pre-period variability, SD=%.1f)\n", sd_ratio_qb_corr2, sd_pre_qb_corr2))

plot(impact2_33b_qc2)

```



```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: causalimpact2_qc33b-c
#| eval: true
#| fig.cap: "Causal impact (Québec only,-3 days pre, local linear trend), controls"
#| fig-width: 9.5
#| fig-height: 5.5
#| fig-dpi: 600
#| results: "hold"

## ── Vector de respuesta ───────────────────────────────────────────────────────
# y33b_lin_qc_c <- c(as.numeric(df33c_qc$sum_velocidad_lead2), rep(NA, 3))
# y33b_lin_qc_c <- log(y33b_lin_qc_c)
lambda_opt <- forecast::BoxCox.lambda(as.numeric(df33c_qc$sum_velocidad_lead2), lower = -2, upper = 2)
y33b_lin_qc_c <- BoxCox(as.numeric(df33c_qc$sum_velocidad_lead2), lambda_opt)
y33b_lin_qc_c <- c(as.numeric(y33b_lin_qc_c), rep(NA, 3))

## ── Estado: tendencia local lineal ───────────────────────────────────────────
ss2bb_qc_c <- list()
ss2bb_qc_c <- AddLocalLinearTrend(ss2bb_qc_c, y33b_lin_qc_c)
ss2bb_qc_c <- AddSeasonal        (ss2bb_qc_c, y33b_lin_qc_c, nseasons = 7, season.duration = 1)
#ss2bb_qc <- AddAutoAr          (ss2bb_qc, y33b_lin_qc, lags = 1)
ss2bb_qc_c <- AddDynamicRegression(ss2bb_qc_c, covariates_full33b_qc)

## ── Ajuste BSTS y CausalImpact ───────────────────────────────────────────────
model33b_qc_c <- bsts(
  y33b_lin_qc_c,
  state.specification = ss2bb_qc_c,
  family = "gaussian",
  niter  = 4e4,
  seed   = 2125
)

impact2_33b_qc_c <- CausalImpact(
  bsts.model           = model33b_qc_c,
  model.args           = list(prior.level.sd = .1, dynamic.regression = TRUE),
  #post.period.response = log(df33c_post_qc$sum_velocidad_lead2[1:3])
  post.period.response = as.numeric(BoxCox(as.numeric(df33c_post_qc$sum_velocidad_lead2[1:3]), lambda_opt))
)

summary(impact2_33b_qc_c)


err_ci_c_qb_corr2 <- impact2_33b_qc_c$series$response - impact2_33b_qc_c$series$point.pred
rmse_ci_c_qb_corr2 <- sqrt(mean(err_ci_c_qb_corr2[1:34]^2))
mean_pre_c_qb_corr2 <- mean(df33c_qc$sum_velocidad_lead2[1:34], na.rm=T) 
sd_pre_c_qb_corr2 <- sd(df33c_qc$sum_velocidad_lead2[1:34], na.rm=T)
nrmse_c_qb_corr2 <- rmse_ci_c_qb_corr2 / mean_pre_c_qb_corr2 # RMSE relativo; 
sd_ratio_c_qb_corr2 <- rmse_ci_c_qb_corr2 / sd_pre_c_qb_corr2 # RMSE relativo a la variabilidad natural

cat(sprintf("RMSE on 35 post-period points: %.2f (average absolute forecast error)\n", rmse_ci_c_qb_corr2))
cat(sprintf("NRMSE: %.2f (RMSE relative to pre-period mean, %.1f)\n", nrmse_c_qb_corr2, mean_pre_c_qb_corr2))
cat(sprintf("SD-ratio: %.2f (RMSE vs pre-period variability, SD=%.1f)\n", sd_ratio_c_qb_corr2, sd_pre_c_qb_corr2))

plot(impact2_33b_qc_c)
```



<br>

# Session info

```{r session-info, echo=T, error=T, message=TRUE, paged.print=TRUE,eval=T, results='hold'}
cat(paste0("R library: ", Sys.getenv("R_LIBS_USER")))
cat(paste0("Date: ",withr::with_locale(new = c('LC_TIME' = 'C'), code =Sys.time())))
cat(paste0("Editor context: ", getwd()))
cat("quarto version: "); system("quarto --version") 

quarto::quarto_version()

save.image("_data/step3.RData")
```

```{r session-info-r, echo=T, error=T, message=TRUE, paged.print=TRUE,eval=T, results='hold'}
#| class-output: center-table

sesion_info <- devtools::session_info()
dplyr::select(
  tibble::as_tibble(sesion_info$packages),
  c(package, loadedversion, source)
) |> 
 knitr::kable(caption = "R packages", format = "html",
      col.names = c("Row number", "Package", "Version"),
    row.names = FALSE,
      align = c("c", "l", "r")) |> 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"),font_size = 12)|> 
  kableExtra::scroll_box(width = "100%", height = "375px")  
```

```{r session-info-python, echo=T, error=T, message=TRUE, paged.print=TRUE,eval=T, results='hold'}
#| class-output: center-table

reticulate::py_list_packages()%>% 
 knitr::kable(caption = "Python packages", format = "html",
      col.names = c("Package", "Version", "Requirement"),
    row.names = FALSE,
      align = c("c", "l", "r", "r"))%>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"),font_size = 12)|>
  kableExtra::scroll_box(width = "100%", height = "375px")  
```


# References