---
title: "Fast followers?/Speed contagion: assessing the impact of Montreal F1 Grand Prix on high-speed ticketing rates (2000-2022)"
description: |
  Step 2. Check analytical strategies
lang: en
date: last-modified #date-format: "D [d]e MMM, YYYY"
bibliography: [_lit/refs.bib]
csl: "_lit/american-medical-association.csl" # (Opcional: estilo de citas, p.ej., APA)
author: "Andrés González Santa Cruz"
format: 
  html:
    css: [_lib/styles.scss]
    code-fold: true
    embed-resources: true
    fig-cap-location: top
lightbox: auto
toc: true
toc-depth: 5
toc-location: left
toc-float: true
editor: source
---

<style>
.scrollable-content {
  max-height: 350px;
  overflow-y: auto;
}
</style>
<style>
pre.scrollable-code {
  max-height: 350px;
  overflow-y: auto;
}
</style>

```{r}
#| message: true
#| include: true
#| warning: true

# remove objects and memory
rm(list=ls());gc()

#remove images
while(!dev.cur())dev.off()
cat("\014")
load(paste0(getwd(),"/_data/step12.RData"))
```

# Load libraries and data

<div class="scrollable-content">
```{r}
#| label: setup
#| message: true
#| warning: true
#| class-output: center-table
#| include: true
#| results: "hold"

#borrar caché
#system("fc-cache -f -v")

#check R version
if(Sys.info()["sysname"]=="Windows"){
if (getRversion() != "4.4.1") { stop("Requiere versión de R 4.4.1. Actual: ", getRversion()) }
}
if(Sys.info()["sysname"]=="Linux"){
if (getRversion() != "4.4.1") { stop("Requiere versión de R 4.4.1. Actual: ", getRversion()) }
}
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:

# install.packages(c("dplyr", #for data
#                    "tidyr", #for data
#                    "lubridate",  #for dates
#                    "openxlsx", #for excel files
#                    "rio", #for importing and exporting data
#                    "purrr", #for iterating in databases
#                    "devtools", #for external packages
#                    "DiagrammeR", #para visualizar DAG
#                    "dagitty",
#                    "ggdag",
#                    "ggplot2" #for graphics
#                    "kableExtra", #pretty tables
#                    "quarto", #for documents
#                    "geosphere" #for coordinates and classifying
#                    "geepack", #for regression
#                    "glmmTMB", #For GLMMs
#                    "DHARMa", #For residual diagnostics
#                    "car", #For hypothesis testing
#                    "brms", #Bayesian model
#                    "bayesplot",
#                    "loo",
#                    "Synth", #for synthetic control method
#                    "weathercan",#for weather data
#                    "sandwich", #cluster robust intervals
#                    "emmeans", #for predictions
#                    "gnm", #Conditional Poisson models
#                    "splines", #nonlinearity 
#                    "geeM", #negative binomial and more flexible GEE models
#                    "PanelMatch", #Matching technique with panel data
#                    "scpi" #control sintético
#                    "nixtlar", #for time series analysis and prediction
#                    "CausalImpact" #for time series causal impact
#                    "forecast" #for time series analysis prediction and decomposition
#                    ))
library(dplyr); library(lubridate); library(tidyverse); library(openxlsx); library(rio); library(purrr); library(dagitty); library(ggdag); library(kableExtra); library(geosphere); library(geepack); library(lme4); library(glmmTMB); library(DHARMa); library(car); library(brms); library(bayesplot); library(loo); library(Synth); library(sandwich); library(emmeans); library(gnm); library(splines); library(geeM); library(plm); library(forecast); library(cowplot); library(bbmle)

#special repository indicated or the package
if(!require(weathercan)){
   install.packages("weathercan", 
                  repos = c("https://ropensci.r-universe.dev", "https://cloud.r-project.org")); library(weathercan)
  }

#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_

if(!require(bpmn)){devtools::install_github("bergant/bpmn")}

#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
sum_dates <- function(x){
 
  cbind.data.frame(
	min= as.Date(min(unclass(as.Date(x)), na.rm=T), origin = "1970-01-01"),
	p001= as.Date(quantile(unclass(as.Date(x)), .001, na.rm=T), origin = "1970-01-01"),
	p005= as.Date(quantile(unclass(as.Date(x)), .005, na.rm=T), origin = "1970-01-01"),
	p025= as.Date(quantile(unclass(as.Date(x)), .025, na.rm=T), origin = "1970-01-01"),
	p25= as.Date(quantile(unclass(as.Date(x)), .25, na.rm=T), origin = "1970-01-01"),
	p50= as.Date(quantile(unclass(as.Date(x)), .5, na.rm=T), origin = "1970-01-01"),
	p75= as.Date(quantile(unclass(as.Date(x)), .75, na.rm=T), origin = "1970-01-01"),
	p975= as.Date(quantile(unclass(as.Date(x)), .975, na.rm=T), origin = "1970-01-01"),
	p995= as.Date(quantile(unclass(as.Date(x)), .995, na.rm=T), origin = "1970-01-01"),
	p999= as.Date(quantile(unclass(as.Date(x)), .999, na.rm=T), origin = "1970-01-01"),
	max= as.Date(max(unclass(as.Date(x)), na.rm=T), origin = "1970-01-01")
  )
}
smd_bin <- function(x,y){
  z <- x*(1-x)
  t <- y*(1-y)
  k <- sum(z,t)
  l <- k/2
  
  return((x-y)/sqrt(l))
  
}

theme_custom_sjplot2 <- function(base_size = 12, base_family = "") {
  theme_minimal(base_size = base_size, base_family = base_family) +
    theme(
      # Text elements
      text = element_text(size = base_size, family = base_family),
      plot.title = element_text(face = "bold", hjust = 0.5, size = base_size * 1.2),
      plot.subtitle = element_text(hjust = 0.5, margin = margin(b = 10)),
      axis.title = element_text(size = base_size, face = "bold"),
      axis.text = element_text(size = base_size * 0.8),
      axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.5),
      axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5),
      axis.title.x = element_text(margin = margin(t = 10)),
      axis.title.y = element_text(margin = margin(r = 10)),
      
      # Plot layout
      plot.margin = margin(t = 20, r = 20, b = 20, l = 20),
      panel.grid.major = element_line(color = "grey80"),
      panel.grid.minor = element_blank(),
      legend.position = "right",
      legend.text = element_text(size = base_size * 0.8),
      legend.title = element_text(size = base_size, face = "bold"),
      legend.background = element_rect(fill = "white", colour = NA),
      legend.box.background = element_rect(colour = "grey80", linetype = "solid"),
      legend.key = element_rect(fill = "white", colour = "white")
    )
}

#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
num_cores <- parallel::detectCores() -1
data.table::setDTthreads(threads = num_cores)#restore_after_fork = NULL, throttle = NULL)

#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#CONFIG #######################################################################
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_

options(scipen=2) #display numbers rather scientific number


#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#CountSynth functions
get_covbal_csnth<- function(countsynth){
  res <- countsynth            # pick any run
  
  dp       <- res$dataprep.main    # Synth-style 'dataprep' object
  W        <- res$unit.weight.full.sample  # CSCM weights (full sample)
  W.scm    <- res$unit.weight.SCM          # Classic SCM weights
  cf.ts    <- res$Estimates                # time-series of fits
  
return(list(
  dataprep = dp,
  W = W,
  W.scm = W.scm,
  cf.ts = cf.ts,
  ATT = res$ATT,
  ATT.scm = res$ATT.scm,
  unit.weight.SCM = res$unit.weight.SCM,
  unit.weight.full.sample = res$unit.weight.full.sample,
  covtab = data.frame(
    variable = rownames(dp$X1),
    treated = as.numeric(dp$X1),
    synthetic = as.numeric(dp$X0 %*% W),
    abs.diff = abs(dp$X1 - dp$X0 %*% W)) |>
    dplyr::mutate(
    ratio = sprintf("%1.2f", treated / synthetic),
    gap_pct_ctrl = sprintf("%1.2f", abs(100 * (treated - synthetic) / synthetic)),
    treated= sprintf("%1.2f", treated),
    synthetic= sprintf("%1.2f", synthetic),
  )
))
}

prepost_rmspe_cntsh<- function(covbal_csnth,t_int){
  # Extract the results from the countSynth object
  cf.ts    <- covbal_csnth$cf.ts
  t_int    <- t_int
  
  # Calculate pre- and post-treatment RMSPE
  max_time <- max(cf.ts$time)
  
  # Create logical vectors for pre- and post-treatment periods
  pre    <- cf.ts$time <  t_int
  post   <- cf.ts$time >= t_int

  rmspe <- function(a,b) sqrt(mean((a-b)^2))
  
  pre.rmspe  <- rmspe(cf.ts$observed[pre], cf.ts$est.counterfactual[pre])
  post.rmspe <- rmspe(cf.ts$observed[post], cf.ts$est.counterfactual[post])
  ratio      <- post.rmspe / pre.rmspe
  return(
    list(
      pre_post_ratio= c(pre=pre.rmspe, post=post.rmspe, ratio=ratio)
      )
  )
}
plot_synth_cscm <- function(res_object, 
                            max_time = NULL, 
                            t_int_man,
                            y_lab = "High-speed rate",
                            colors = c("SCM" = "black", "CSCM" = "red", "Obs" = "gray60")) {
  
  # Validate inputs
  if (!inherits(res_object, "list")) stop("res_object must be a list")
  if (!all(c("dataprep.main", "unit.weight.SCM", "unit.weight.full.sample") %in% names(res_object))) {
    stop("res_object missing required components")
  }
  if (is.null(max_time)) max_time <- nrow(res_object$dataprep.main$Y0plot)
  
  # Calculate counterfactuals
  synth.orig <- res_object$dataprep.main$Y0plot %*% as.numeric(res_object$unit.weight.SCM)
  cscm.cf <- res_object$dataprep.main$Y0plot %*% as.numeric(res_object$unit.weight.full.sample)
  year <- 1:max_time
  
  # Create data frames
  adf <- data.frame(Year = year, Y = synth.orig[1:max_time], Method = "SCM")
  bdf <- data.frame(Year = year, Y = cscm.cf[1:max_time], Method = "CSCM")
  cdf <- data.frame(Year = year, Y = res_object$dataprep.main$Y1plot[1:max_time], Method = "Obs")
  
  # Combine data
  cfdf <- rbind(adf, bdf, cdf)
  
  # Create plot
  ggplot(cfdf, aes(y = Y, x = Year, group = Method, linetype = Method, color = Method)) + 
    geom_vline(xintercept = t_int_man, linetype = "dashed") +
    geom_line(linewidth = 1.2) +
    scale_color_manual(values = colors) +
    theme_bw() +
    labs(y = y_lab, x = "") +
    theme(
      text = element_text(size = 15),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.line = element_line(colour = "black"),
      legend.position = "bottom"
    # Ensure linetype scale if needed (uncomment if necessary)
    # scale_linetype_manual(values = c("SCM" = "solid", "CSCM" = "solid", "Observed" = "solid"))
    )
}


tidy_exp_gnm <- function(model, conf.level = 0.95, use_t = FALSE) {
  td <- broom::tidy(model)
  alpha <- 1 - conf.level
  crit <- if (use_t) {
    stats::qt(1 - alpha/2, df = model$df.residual)
  } else {
    stats::qnorm(1 - alpha/2)
  }
    td %>%
    mutate(
      conf.low_log  = estimate - crit * std.error,
      conf.high_log = estimate + crit * std.error,
      rate_ratio    = exp(estimate),
      rr_conf.low   = exp(conf.low_log),
      rr_conf.high  = exp(conf.high_log)
    ) %>%
    dplyr::select(term, rate_ratio, rr_conf.low, rr_conf.high, std.error, p.value) |> 
    data.frame() |> 
    dplyr::slice(1)
}

tidy_exp_gnm_corr <- function(modelo) {
    as_tibble(Epi::ci.exp(modelo, subset = 1, pval = TRUE),
              rownames = "term")
}
```


</div>

  
```{r}
#| message: true
#| include: true
#| warning: false
#| error: false
#| eval: false
#| fig.show: "hold"
#| fig.align: "center"
#| paged.print: true
#| fig.cap: "Directed acyclic graph"
#| results: "hold"
#| fig-width: 9.5
#| fig-height: 5.5
#| fig-dpi: 600
#| label: dag
#| fig-show: hide

#-- Fórmula 1 – ventana de riesgo
#-- Accidentes por exceso de velocidad
#--Weather (temperatura, lluvia…)
#-- Day-of-week / season / calendario
#-- Rasgos fijos de la comuna (infraestructura, cultura vial)
dag <- dagitty('
dag {
  F1   [exposure,   pos="-2.000,  0.000"]
  Y    [outcome,    pos=" 2.000,  0.000"]
  W    [adjusted,   pos="-1.000,  3.250"]
  DOW  [adjusted,   pos="-1.000, -3.250"]
  U    [unobserved, pos=" -3.000,  2.000"]

  F1  -> Y
  W   -> F1
  W   -> Y
  DOW -> F1
  DOW -> Y
  U   -> Y
}
')
# coords <- data.frame(
#   x = c(-2,  2, -1, -1,  3),   # ← exactly the positions you want
#   y = c( 0,  0,  3, -3,  0),
#   row.names = c("F1", "Y", "W", "DOW", "U")
# )
# 
# coordinates(dag) <- coords     # fix positions inside dagitty

# data
tidy  <- tidy_dagitty(dag)

edges <- tidy %>%                      # sólo flechas →
  filter(direction == "->") %>%
  mutate(
    edge_colour    = if_else(name == "U", "grey60", "black"),
    edge_linetype  = if_else(name == "U", "dashed",  "solid"),
    edge_linewidth = 0.9                               # grosor
  )

nodes <- tidy %>% data.frame() |>  distinct(name, x, y) %>%             # nodos
  mutate(shape = if_else(name %in% c("W","DOW"),
                         "square", "circle"))

## 3 · Plot -------------------------------------------------------------
ggplot() +
  geom_dag_edges_link(                     # ← solo flechas →
    data = edges,
    aes(x = x, y = y, xend = xend, yend = yend,
        edge_colour    = edge_colour,
        edge_linetype  = edge_linetype,
        edge_linewidth = edge_linewidth)
  ) +
  geom_dag_node(
    data = nodes,
    aes(x = x, y = y, shape = shape),
    fill   = "white", colour = "black", size = 18
  ) +
  geom_dag_text(data = nodes, aes(x = x, y = y, label = name)) +
  scale_shape_manual(values = c(square = 15, circle = 16)) + 
  #scale_edge_colour_identity() +
  #scale_edge_linetype_identity() +
  #edge_width_identity() +   # usa edge_width_identity() si tu ggraph ≤ 2.0
  theme_void()+
  theme(legend.position = "none")

```

<br>


```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: test-association-autorregressive
#| eval: false
#| results: "hold"
library(DHARMa)
#simulationOutput <- simulateResiduals(fittedModel = model_did1, plot = TRUE)

cities <- unique(collisions_weather_corr$cluster)
for (c in cities) {
  cat("\nEvaluando autocorrelación para cluster:", c, "\n")
  
  city_data <- collisions_weather_corr %>% filter(cluster == c) %>% arrange(iso_yday)
  
  indices <- which(collisions_weather_corr$cluster == c)
  city_residuals <- simulationOutput$scaledResiduals[indices]
  
  # Verificar que city_data y city_residuals tengan la misma longitud
  if(length(city_residuals) != nrow(city_data)) {
    cat("Error: longitud de residuos y datos no coincide para cluster:", c, "\n")
    next
  }
  
  # Verificar que no haya NA en residuos ni en fechas
  if(any(is.na(city_residuals)) || any(is.na(city_data$date))) {
    cat("Error: NA en residuos o fechas para cluster:", c, "\n")
    next
  }
  
  # Ordenar residuos según fechas
  order_idx <- order(city_data$iso_yday)
  city_data <- city_data[order_idx, ]
  city_residuals <- city_residuals[order_idx]
  
  print(paste("Length residuos:", length(city_residuals)))
  print(paste("Length fechas:", length(city_data$date)))
  print(paste("Duplicados fechas:", any(duplicated(city_data$date))))
  print(paste("NA en residuos:", sum(is.na(city_residuals))))
  
  # Verificar que fechas estén ordenadas y sin duplicados
  if(any(duplicated(city_data$date))) {
    cat("Error: fechas duplicadas para cluster:", c, "\n")
    next
  }
  
  # Finalmente, llamar a la función
  tryCatch({
    testTemporalAutocorrelation(city_residuals, time = city_data$date)
    mtext(text = paste("Autocorr, cluster:", c), side = 3, line = 3, cex = 1.2, font = 2)
  }, error = function(e) {
    cat("Error en testTemporalAutocorrelation para cluster:", c, "\n", e$message, "\n")
  })
}

invisible("No hay evidencia significativa de autocorrelación temporal en los residuos del modelo para ninguna de las ciudades.")


# simulationOutput <- simulateResiduals(fittedModel = model_final, plot = TRUE)
# testDispersion(simulationOutput)
# testZeroInflation(simulationOutput)
```


### GNM


```{r}
# Case-crossover design
# 
# Methods:
# Only cases: cardiovascular defects 3870
# 
# Exposure:
# Case window: trimethoprim use any time during the second
# 	and third months after the last menstrual period,
# Reference window: use in 2 months prior to last menstrual
# 	period within mothers of case infants.
# Secondary reference windows: and 2 months prior and 2
# 	months after the primary reference window
# 
# Analysis:
# 

#https://docs.google.com/document/d/1shhxi_PEkgc8N4riqV4lOowPcGPo2mdj4-PGA779oUk/edit?tab=t.0
#https://drive.google.com/drive/u/0/folders/1gwdKKOVZd2JEHmb0oXt34FSpvea1oH7G no aplica tanto potque el análisis es por persona, es Conditional logistic regression y ve mortalidad. se busca para esa misma persona variables como días de tiempo

#https://docs.google.com/presentation/d/1Khvb5C8ZTaw3HKfvx3oAKD7fBZHPcd_v/edit?slide=id.p42#slide=id.p42
#https://docs.google.com/spreadsheets/d/1znDbWVXO2A5uSKiVazQwFJKZcrrH-VZXnQsG-INBQ-w/edit?gid=0#gid=0
#https://drive.google.com/drive/u/0/folders/1TuEbUzsFeLtQtiuP3a6uSw3DulIq-_LV #tópicos avanzados en Epi

```

We stratify by control states, according to the **meeting** with José Ignacio.

```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: test-association-gnm
#| eval: true
#| results: "hold"
#Factor variable that identifies each case-control block
#Ej.: interaction(ciudad, mes, día_semana) o id_persona en estudios de eventos agudos.

## ── Helper para φ  (sobre‑dispersión) ──────────────────────────────────────────
get_phi <- function(mod) {
  max( sum(residuals(mod, type = "pearson")^2) / df.residual(mod), 1 )  # ≤1 ⇒ 1
}

## ── Función QIC para modelos gnm quasi‑Poisson ────────────────────────────────
qic_gnm <- function(mod_quasi) {
  ## 1)  φ desde el quasi
  phi <- get_phi(mod_quasi)

  ## 2)  refit Poisson (mismo call)  → logLik disponible
  mod_pois <- update(mod_quasi, family = poisson)

  ## 3)  qAIC = -2·logLik/φ + 2·k    (bbmle lo hace por ti)
  bbmle::qAIC(mod_pois, dispersion = phi)
}



cat("Treated (w/o Sherbrooke), contemporaneous effects\n")
model_gnm_tratados <- gnm(
  sum_velocidad_lead1 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Post",exposure_window) & treated==TRUE)
)
summary(model_gnm_tratados)


cat("Controls (w/o Sherbrooke), contemporaneous effects\n")
model_gnm_controles <- gnm(
  sum_velocidad_lead1 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Post",exposure_window) & treated==FALSE)
)
summary(model_gnm_controles)

b1  <- model_gnm_controles$coefficients[[1]]; se1 <- sqrt(diag(vcov(model_gnm_controles)))[[1]]
b2  <- model_gnm_tratados$coefficients[[1]]; se2 <- sqrt(diag(vcov(model_gnm_tratados)))[[1]]
z   <- (b1 - b2) / sqrt(se1^2 + se2^2)
p   <- 2 * pnorm(-abs(z))     # 0.32
cat("Controls vs. Treated (w/o Sherbrooke), contemporaneous effects;  p-values\n")
p


log_rrr  <- b2- b1
se_rrr   <- sqrt(se1^2 + se2^2)   # var(Δ) = var1 + var2

## RRR y IC 95 %
RRR      <- exp(log_rrr)
lo95     <- exp(log_rrr - 1.96 * se_rrr)
hi95     <- exp(log_rrr + 1.96 * se_rrr)

cat(sprintf("RRR = %.3f (95%% CI %.3f–%.3f)\n", RRR, lo95, hi95))



cat("Treated (w/o Sherbrooke), effects after 1 day\n")
model_gnm_tratados_lead1 <- gnm(
  sum_velocidad_lead1 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Exp",exposure_window) & treated==TRUE)
)
summary(model_gnm_tratados_lead1)


cat("Controls (w/o Sherbrooke), effects after 1 day\n")
model_gnm_controles_lead1 <- gnm(
  sum_velocidad_lead1 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Exp",exposure_window) & treated==FALSE)
)
summary(model_gnm_controles_lead1)

b1_lead1  <- model_gnm_controles_lead1$coefficients[[1]]; se1_lead1 <- sqrt(diag(vcov(model_gnm_controles_lead1)))[[1]]
b2_lead1  <- model_gnm_tratados_lead1$coefficients[[1]]; se2_lead1 <- sqrt(diag(vcov(model_gnm_tratados_lead1)))[[1]]
z_lead1   <- (b1_lead1 - b2_lead1) / sqrt(se1_lead1^2 + se2_lead1^2)
p_lead1   <- 2 * pnorm(-abs(z_lead1))     # 0.32
cat("Controls vs. Treated (w/o Sherbrooke), effects after 1 day;  p-values\n")
p_lead1


log_rrr  <- b2_lead1- b1_lead1
se_rrr   <- sqrt(se1_lead1^2 + se2_lead1^2)   # var(Δ) = var1 + var2

## RRR y IC 95 %
RRR_lead1      <- exp(log_rrr)
lo95_lead1     <- exp(log_rrr - 1.96 * se_rrr)
hi95_lead1     <- exp(log_rrr + 1.96 * se_rrr)

cat(sprintf("RRR = %.3f (95%% CI %.3f–%.3f)\n", RRR_lead1, lo95_lead1, hi95_lead1))



cat("Treated (w/o Sherbrooke), effects after 2 days\n")
model_gnm_tratados_lead2 <- gnm(
  sum_velocidad_lead2 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Exp",exposure_window) & treated==TRUE)
)
summary(model_gnm_tratados_lead2)


cat("Controls (w/o Sherbrooke), effects after 2 days\n")
model_gnm_controles_lead2 <- gnm(
  sum_velocidad_lead2 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Exp",exposure_window) & treated==FALSE)
)
summary(model_gnm_controles_lead2)

b1_lead2  <- model_gnm_controles_lead2$coefficients[[1]]; se1_lead2 <- sqrt(diag(vcov(model_gnm_controles_lead2)))[[1]]
b2_lead2  <- model_gnm_tratados_lead2$coefficients[[1]]; se2_lead2 <- sqrt(diag(vcov(model_gnm_tratados_lead2)))[[1]]
z_lead2   <- (b1_lead2 - b2_lead2) / sqrt(se1_lead2^2 + se2_lead2^2)
p_lead2   <- 2 * pnorm(-abs(z_lead2))     # 0.32
cat("Controls vs. Treated (w/o Sherbrooke), effects after 2 days;  p-values\n")
p_lead2


log_rrr  <- b2_lead2- b1_lead2
se_rrr   <- sqrt(se1_lead2^2 + se2_lead2^2)   # var(Δ) = var1 + var2

## RRR y IC 95 %
RRR_lead2      <- exp(log_rrr)
lo95_lead2     <- exp(log_rrr - 1.96 * se_rrr)
hi95_lead2     <- exp(log_rrr + 1.96 * se_rrr)

cat(sprintf("RRR = %.3f (95%% CI %.3f–%.3f)\n", RRR_lead2, lo95_lead2, hi95_lead2))


#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
cat("With Sherbrooke\n")

cat("Treated (w/ Sherbrooke), contemporaneous effects\n")
model_gnm_tratados_sens <- gnm(
  sum_velocidad_lead1 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Exp",exposure_window) & treated_sens==TRUE)
)
summary(model_gnm_tratados_sens)


cat("Controls (w/ Sherbrooke), contemporaneous effects\n")
model_gnm_controles_sens <- gnm(
  sum_velocidad_lead1 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Exp",exposure_window) & treated_sens==FALSE)
)
summary(model_gnm_controles_sens)

b1_sens  <- model_gnm_controles_sens$coefficients[[1]]; se1_sens <- sqrt(diag(vcov(model_gnm_controles_sens)))[[1]]
b2_sens  <- model_gnm_tratados_sens$coefficients[[1]]; se2_sens <- sqrt(diag(vcov(model_gnm_tratados_sens)))[[1]]
z_sens   <- (b1_sens - b2_sens) / sqrt(se1_sens^2 + se2_sens^2)
p_sens   <- 2 * pnorm(-abs(z_sens))     # 0.32
cat("Controls (w/ Sherbrooke) vs. Treated, contemporaneous effects;  p-values\n")
p_sens


log_rrr  <- b2_sens- b1_sens 
se_rrr   <- sqrt(se1_sens^2 + se2_sens^2)   # var(Δ) = var1 + var2

## RRR y IC 95 %
RRR_sens      <- exp(log_rrr)
lo95_sens     <- exp(log_rrr - 1.96 * se_rrr)
hi95_sens     <- exp(log_rrr + 1.96 * se_rrr)

cat(sprintf("RRR = %.3f (95%% CI %.3f–%.3f)\n", RRR_sens, lo95_sens, hi95_sens))


cat("Treated (w/ Sherbrooke), effects after 1 day\n")
model_gnm_tratados_lead1_sens <- gnm(
  sum_velocidad_lead1 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Exp",exposure_window) & treated_sens==TRUE)
)
summary(model_gnm_tratados_lead1_sens)


cat("Controls (w/ Sherbrooke), effects after 1 day\n")
model_gnm_controles_lead1_sens <- gnm(
  sum_velocidad_lead1 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Exp",exposure_window) & treated_sens==FALSE)
)
summary(model_gnm_controles_lead1_sens)

b1_lead1_sens  <- model_gnm_controles_lead1_sens$coefficients[[1]]; se1_lead1_sens <- sqrt(diag(vcov(model_gnm_controles_lead1_sens)))[[1]]
b2_lead1_sens  <- model_gnm_tratados_lead1_sens$coefficients[[1]]; se2_lead1_sens <- sqrt(diag(vcov(model_gnm_tratados_lead1_sens)))[[1]]
z_lead1_sens   <- (b1_lead1_sens - b2_lead1_sens) / sqrt(se1_lead1_sens^2 + se2_lead1_sens^2)
p_lead1_sens   <- 2 * pnorm(-abs(z_lead1_sens))     # 0.32
cat("Controls (w/ Sherbrooke) vs. Treated, effects after 1 day;  p-values\n")
p_lead1_sens

log_rrr  <- b2_lead1_sens- b1_lead1_sens
se_rrr   <- sqrt(se1_lead1_sens^2 + se2_lead1_sens^2)   # var(Δ) = var1 + var2

## RRR y IC 95 %
RRR_lead1_sens      <- exp(log_rrr)
lo95_lead1_sens     <- exp(log_rrr - 1.96 * se_rrr)
hi95_lead1_sens     <- exp(log_rrr + 1.96 * se_rrr)

cat(sprintf("RRR = %.3f (95%% CI %.3f–%.3f)\n", RRR_lead1_sens, lo95_lead1_sens, hi95_lead1_sens))


cat("Treated (w/ Sherbrooke), effects after 2 days\n")
model_gnm_tratados_lead2_sens <- gnm(
  sum_velocidad_lead2 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Exp",exposure_window) & treated_sens==TRUE)
)
summary(model_gnm_tratados_lead2_sens)


cat("Controls (w/ Sherbrooke), effects after 2 days\n")
model_gnm_controles_lead2_sens <- gnm(
  sum_velocidad_lead2 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Exp",exposure_window) & treated_sens==FALSE)
)
summary(model_gnm_controles_lead2_sens)

b1_lead2_sens  <- model_gnm_controles_lead2_sens$coefficients[[1]]; se1_lead2_sens <- sqrt(diag(vcov(model_gnm_controles_lead2_sens)))[[1]]
b2_lead2_sens  <- model_gnm_tratados_lead2_sens$coefficients[[1]]; se2_lead2_sens <- sqrt(diag(vcov(model_gnm_tratados_lead2_sens)))[[1]]
z_lead2_sens   <- (b1_lead2_sens - b2_lead2_sens) / sqrt(se1_lead2_sens^2 + se2_lead2_sens^2)
p_lead2_sens   <- 2 * pnorm(-abs(z_lead2_sens))     # 0.32
cat("Controls (w/ Sherbrooke) vs. Treated, effects after 2 days;  p-values\n")
p_lead2_sens


log_rrr  <- b2_lead2_sens-b1_lead2_sens
se_rrr   <- sqrt(se1_lead2_sens^2 + se2_lead2_sens^2)   # var(Δ) = var1 + var2

## RRR y IC 95 %
RRR_lead2_sens      <- exp(log_rrr)
lo95_lead2_sens     <- exp(log_rrr - 1.96 * se_rrr)
hi95_lead2_sens     <- exp(log_rrr + 1.96 * se_rrr)

cat(sprintf("RRR = %.3f (95%% CI %.3f–%.3f)\n", RRR_lead2_sens, lo95_lead2_sens, hi95_lead2_sens))

qic_gnm_tratados             <- qic_gnm(model_gnm_tratados)
qic_gnm_tratados_lead1       <- qic_gnm(model_gnm_tratados_lead1)
qic_gnm_tratados_lead2       <- qic_gnm(model_gnm_tratados_lead2)
qic_gnm_tratados_sens        <- qic_gnm(model_gnm_tratados_sens)
qic_gnm_tratados_lead1_sens  <- qic_gnm(model_gnm_tratados_lead1_sens)
qic_gnm_tratados_lead2_sens  <- qic_gnm(model_gnm_tratados_lead2_sens)

qic_gnm_controles            <- qic_gnm(model_gnm_controles)
qic_gnm_controles_lead1      <- qic_gnm(model_gnm_controles_lead1)
qic_gnm_controles_lead2      <- qic_gnm(model_gnm_controles_lead2)
qic_gnm_controles_sens       <- qic_gnm(model_gnm_controles_sens)
qic_gnm_controles_lead1_sens <- qic_gnm(model_gnm_controles_lead1_sens)
qic_gnm_controles_lead2_sens <- qic_gnm(model_gnm_controles_lead2_sens)


## ─── 2. Vector with names ───────────────────────────────────────────
qic_vec <- c(
  "Treated (base)"            = qic_gnm_tratados,
  "Treated + lead 1"          = qic_gnm_tratados_lead1,
  "Treated + lead 2"          = qic_gnm_tratados_lead2,
  "Treated sensitivity (- Sherbrooke)"      = qic_gnm_tratados_sens,
  "Treated sens + lead 1"     = qic_gnm_tratados_lead1_sens,
  "Treated sens + lead 2"     = qic_gnm_tratados_lead2_sens,
  "Controls (base)"           = qic_gnm_controles,
  "Controls + lead 1"         = qic_gnm_controles_lead1,
  "Controls + lead 2"         = qic_gnm_controles_lead2,
  "Controls sensitivity (+ Sherbrooke)"     = qic_gnm_controles_sens,
  "Controls sens + lead 1"    = qic_gnm_controles_lead1_sens,
  "Controls sens + lead 2"    = qic_gnm_controles_lead2_sens
)

## ─── 3. print results ───────────────────────────────────────
cat(
  "\n≡ Comparison of models through QIC (quasi‑AIC) ≡",
  "\n(‑2·ℓ / φ  +  2·k ;  lower ⇒ better)\n\n",
  paste(names(qic_vec), sprintf(": %.1f", qic_vec), collapse = "\n"),
  "\n\n≡ Ranking (better → worse) ≡\n",
  paste(names(sort(qic_vec)), sprintf(": %.1f", sort(qic_vec)), collapse = "\n"),
  "\n"
)

```

We made placebo tests with post-exposure.

```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: test-association-gnm-placebo
#| eval: true
#| results: "hold"

cat("Treated (w/o Sherbrooke), contemporaneous effects\n")
placebo_model_gnm_tratados <- gnm(
  sum_velocidad_lead1 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Post",exposure_window) & treated==TRUE)
)
summary(placebo_model_gnm_tratados)


cat("Controls (w/o Sherbrooke), contemporaneous effects\n")
placebo_model_gnm_controles <- gnm(
  sum_velocidad_lead1 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Post",exposure_window) & treated==FALSE)
)
summary(placebo_model_gnm_controles)

pb1  <- placebo_model_gnm_controles$coefficients[[1]]; pse1 <- sqrt(diag(vcov(placebo_model_gnm_controles)))[[1]]
pb2  <- placebo_model_gnm_tratados$coefficients[[1]]; pse2 <- sqrt(diag(vcov(placebo_model_gnm_tratados)))[[1]]
pz   <- (pb1 - pb2) / sqrt(pse1^2 + pse2^2)
pp   <- 2 * pnorm(-abs(pz))     # 0.32
cat("Controls vs. Treated (w/o Sherbrooke), contemporaneous effects;  p-values\n")
pp

log_rrr  <- pb2-pb1
se_rrr   <- sqrt(pse1^2 + pse2^2)   # var(Δ) = var1 + var2

## RRR y IC 95 %
RRR_p      <- exp(log_rrr)
lo95_p     <- exp(log_rrr - 1.96 * se_rrr)
hi95_p     <- exp(log_rrr + 1.96 * se_rrr)

cat(sprintf("RRR = %.3f (95%% CI %.3f–%.3f)\n", RRR_p, lo95_p, hi95_p))


cat("Treated (w/o Sherbrooke), effects after 1 day\n")
placebo_model_gnm_tratados_lead1 <- gnm(
  sum_velocidad_lead1 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Post",exposure_window) & treated==TRUE)
)
summary(placebo_model_gnm_tratados_lead1)


cat("Controls (w/o Sherbrooke), effects after 1 day\n")
placebo_model_gnm_controles_lead1 <- gnm(
  sum_velocidad_lead1 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Post",exposure_window) & treated==FALSE)
)
summary(placebo_model_gnm_controles_lead1)

pb1_lead1  <- placebo_model_gnm_controles_lead1$coefficients[[1]]; pse1_lead1 <- sqrt(diag(vcov(placebo_model_gnm_controles_lead1)))[[1]]
pb2_lead1  <- placebo_model_gnm_tratados_lead1$coefficients[[1]]; pse2_lead1 <- sqrt(diag(vcov(placebo_model_gnm_tratados_lead1)))[[1]]
pz_lead1   <- (pb1_lead1 - pb2_lead1) / sqrt(pse1_lead1^2 + pse2_lead1^2)
pp_lead1   <- 2 * pnorm(-abs(z_lead1))     # 0.32
cat("Controls vs. Treated (w/o Sherbrooke), effects after 1 day;  p-values\n")
pp_lead1


log_rrr  <- pb2_lead1-pb1_lead1 
se_rrr   <- sqrt(pse1_lead1^2 + pse2_lead1^2)   # var(Δ) = var1 + var2

## RRR y IC 95 %
RRR_p_lead1      <- exp(log_rrr)
lo95_p_lead1     <- exp(log_rrr - 1.96 * se_rrr)
hi95_p_lead1     <- exp(log_rrr + 1.96 * se_rrr)

cat(sprintf("RRR = %.3f (95%% CI %.3f–%.3f)\n", RRR_p_lead1, lo95_p_lead1, hi95_p_lead1))

cat("Treated (w/o Sherbrooke), effects after 2 days\n")
placebo_model_gnm_tratados_lead2 <- gnm(
  sum_velocidad_lead2 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Post",exposure_window) & treated==TRUE)
)
summary(placebo_model_gnm_tratados_lead2)


cat("Controls (w/o Sherbrooke), effects after 2 days\n")
placebo_model_gnm_controles_lead2 <- gnm(
  sum_velocidad_lead2 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Post",exposure_window) & treated==FALSE)
)
summary(placebo_model_gnm_controles_lead2)

pb1_lead2  <- placebo_model_gnm_controles_lead2$coefficients[[1]]; pse1_lead2 <- sqrt(diag(vcov(placebo_model_gnm_controles_lead2)))[[1]]
pb2_lead2  <- placebo_model_gnm_tratados_lead2$coefficients[[1]]; pse2_lead2 <- sqrt(diag(vcov(placebo_model_gnm_tratados_lead2)))[[1]]
pz_lead2   <- (pb1_lead2 - pb2_lead2) / sqrt(pse1_lead2^2 + pse2_lead2^2)
pp_lead2   <- 2 * pnorm(-abs(pz_lead2))     # 0.32
cat("Controls vs. Treated (w/o Sherbrooke), effects after 2 days;  p-values\n")
pp_lead2



log_rrr  <- pb2_lead2- pb1_lead2
se_rrr   <- sqrt(pse1_lead2^2 + pse2_lead2^2)   # var(Δ) = var1 + var2

## RRR y IC 95 %
RRR_p_lead2      <- exp(log_rrr)
lo95_p_lead2     <- exp(log_rrr - 1.96 * se_rrr)
hi95_p_lead2     <- exp(log_rrr + 1.96 * se_rrr)

cat(sprintf("RRR = %.3f (95%% CI %.3f–%.3f)\n", RRR_p_lead2, lo95_p_lead2, hi95_p_lead2))


#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
cat("With Sherbrooke\n")

cat("Treated (w/ Sherbrooke), contemporaneous effects\n")
placebo_model_gnm_tratados_sens <- gnm(
  sum_velocidad_lead1 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Post",exposure_window) & treated_sens==TRUE)
)
summary(placebo_model_gnm_tratados_sens)


cat("Controls (w/ Sherbrooke), contemporaneous effects\n")
placebo_model_gnm_controles_sens <- gnm(
  sum_velocidad_lead1 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Post",exposure_window) & treated_sens==FALSE)
)
summary(placebo_model_gnm_controles_sens)

pb1_sens  <- placebo_model_gnm_controles_sens$coefficients[[1]]; pse1_sens <- sqrt(diag(vcov(placebo_model_gnm_controles_sens)))[[1]]
pb2_sens  <- placebo_model_gnm_tratados_sens$coefficients[[1]]; pse2_sens <- sqrt(diag(vcov(placebo_model_gnm_tratados_sens)))[[1]]
pz_sens   <- (pb1_sens - pb2_sens) / sqrt(pse1_sens^2 + pse2_sens^2)
pp_sens   <- 2 * pnorm(-abs(pz_sens))     # 0.32
cat("Controls (w/ Sherbrooke) vs. Treated, contemporaneous effects;  p-values\n")
pp_sens


log_rrr  <- pb2_sens-pb1_sens
se_rrr   <- sqrt(pse1_sens^2 + pse2_sens^2)   # var(Δ) = var1 + var2

## RRR y IC 95 %
RRR_p_sens      <- exp(log_rrr)
lo95_p_sens     <- exp(log_rrr - 1.96 * se_rrr)
hi95_p_sens     <- exp(log_rrr + 1.96 * se_rrr)

cat(sprintf("RRR = %.3f (95%% CI %.3f–%.3f)\n", RRR_p_sens, lo95_p_sens, hi95_p_sens))

cat("Treated (w/ Sherbrooke), effects after 1 day\n")
placebo_model_gnm_tratados_lead1_sens <- gnm(
  sum_velocidad_lead1 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Post",exposure_window) & treated_sens==TRUE)
)
summary(placebo_model_gnm_tratados_lead1_sens)


cat("Controls (w/ Sherbrooke), effects after 1 day\n")
placebo_model_gnm_controles_lead1_sens <- gnm(
  sum_velocidad_lead1 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Post",exposure_window) & treated_sens==FALSE)
)
summary(placebo_model_gnm_controles_lead1_sens)

pb1_lead1_sens  <- placebo_model_gnm_controles_lead1_sens$coefficients[[1]]; pse1_lead1_sens <- sqrt(diag(vcov(placebo_model_gnm_controles_lead1_sens)))[[1]]
pb2_lead1_sens  <- placebo_model_gnm_tratados_lead1_sens$coefficients[[1]]; pse2_lead1_sens <- sqrt(diag(vcov(placebo_model_gnm_tratados_lead1_sens)))[[1]]
pz_lead1_sens   <- (pb1_lead1_sens - pb2_lead1_sens) / sqrt(pse1_lead1_sens^2 + pse2_lead1_sens^2)
pp_lead1_sens   <- 2 * pnorm(-abs(pz_lead1_sens))     # 0.32
cat("Controls (w/ Sherbrooke) vs. Treated, effects after 1 day;  p-values\n")
pp_lead1_sens


log_rrr  <- pb2_lead1_sens-pb1_lead1_sens
se_rrr   <- sqrt(pse1_lead1_sens^2 + pse2_lead1_sens^2)   # var(Δ) = var1 + var2

## RRR y IC 95 %
RRR_p_lead1_sens      <- exp(log_rrr)
lo95_p_lead1_sens     <- exp(log_rrr - 1.96 * se_rrr)
hi95_p_lead1_sens     <- exp(log_rrr + 1.96 * se_rrr)

cat(sprintf("RRR = %.3f (95%% CI %.3f–%.3f)\n", RRR_p_lead1_sens, lo95_p_lead1_sens, hi95_p_lead1_sens))

cat("Treated (w/ Sherbrooke), effects after 2 days\n")
placebo_model_gnm_tratados_lead2_sens <- gnm(
  sum_velocidad_lead2 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Post",exposure_window) & treated_sens==TRUE)
)
summary(placebo_model_gnm_tratados_lead2_sens)


cat("Controls (w/ Sherbrooke), effects after 2 days\n")
placebo_model_gnm_controles_lead2_sens <- gnm(
  sum_velocidad_lead2 ~ relevel(factor(exposure_window), ref="Pre-exposure")+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
    offset(off_lic),
  eliminate = cluster,                            # <── “conditional” part
  family    = quasipoisson(link = "log"),
  data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Post",exposure_window) & treated_sens==FALSE)
)
summary(placebo_model_gnm_controles_lead2_sens)

pb1_lead2_sens  <- placebo_model_gnm_controles_lead2_sens$coefficients[[1]]; pse1_lead2_sens <- sqrt(diag(vcov(placebo_model_gnm_controles_lead2_sens)))[[1]]
pb2_lead2_sens  <- placebo_model_gnm_tratados_lead2_sens$coefficients[[1]]; pse2_lead2_sens <- sqrt(diag(vcov(placebo_model_gnm_tratados_lead2_sens)))[[1]]
pz_lead2_sens   <- (pb1_lead2_sens - pb2_lead2_sens) / sqrt(pse1_lead2_sens^2 + pse2_lead2_sens^2)
pp_lead2_sens   <- 2 * pnorm(-abs(pz_lead2_sens))     # 0.32
cat("Controls (w/ Sherbrooke) vs. Treated, effects after 2 days;  p-values\n")
pp_lead2_sens

log_rrr  <- pb2_lead2_sens-pb1_lead2_sens
se_rrr   <- sqrt(pse1_lead2_sens^2 + pse2_lead2_sens^2)   # var(Δ) = var1 + var2

## RRR y IC 95 %
RRR_p_lead2_sens      <- exp(log_rrr)
lo95_p_lead2_sens     <- exp(log_rrr - 1.96 * se_rrr)
hi95_p_lead2_sens     <- exp(log_rrr + 1.96 * se_rrr)

cat(sprintf("RRR = %.3f (95%% CI %.3f–%.3f)\n", RRR_p_lead2_sens, lo95_p_lead2_sens, hi95_p_lead2_sens))


cbind.data.frame(type=c("Main","","", "", "Placebo","","",""),
rbind.data.frame(
                tidy_exp_gnm_corr(model_gnm_tratados_sens),
                tidy_exp_gnm_corr(model_gnm_controles_sens),
                tidy_exp_gnm_corr(model_gnm_tratados),
                tidy_exp_gnm_corr(model_gnm_controles),
                tidy_exp_gnm_corr(placebo_model_gnm_tratados_sens),
                tidy_exp_gnm_corr(placebo_model_gnm_controles_sens),                 
                tidy_exp_gnm_corr(placebo_model_gnm_tratados),
                tidy_exp_gnm_corr(placebo_model_gnm_controles)),
  sig= round(c(p_sens, NA, p, NA, pp_sens, NA, pp, NA),3),
  RRR= c(sprintf("%.2f (%.2f–%.2f)\n", RRR_sens, lo95_sens, hi95_sens), NA, sprintf("%.2f (%.2f–%.2f)\n", RRR, lo95, hi95), NA, sprintf("%.2f (%.2f–%.2f)\n", RRR_p_sens, lo95_p_sens, hi95_p_sens), NA, sprintf("%.2f (%.2f–%.2f)\n", RRR_p, lo95_p, hi95_p), NA))|> 
  mutate(RR= sprintf("%.2f (%.2f–%.2f)", `exp(Est.)`, `2.5%`, `97.5%`))|>
  mutate(sig= sprintf("%.3f", sig))|> 
  rename("p.value"="P")|> 
  dplyr::select(term, p.value, RR, sig, RRR)|> 
  knitr::kable("markdown", caption="Main, high-speed collisions, instantaneous effect", digits=3, na= "-")

cbind.data.frame(type=c("Main","","", "", "Placebo","","",""),
rbind.data.frame(
                tidy_exp_gnm_corr(model_gnm_tratados_lead1_sens),
                tidy_exp_gnm_corr(model_gnm_controles_lead1_sens),
                tidy_exp_gnm_corr(model_gnm_tratados_lead1),
                tidy_exp_gnm_corr(model_gnm_controles_lead1),
                tidy_exp_gnm_corr(placebo_model_gnm_tratados_lead1_sens),
                tidy_exp_gnm_corr(placebo_model_gnm_controles_lead1_sens),
                tidy_exp_gnm_corr(placebo_model_gnm_tratados_lead1),
                tidy_exp_gnm_corr(placebo_model_gnm_controles_lead1)),
sig= round(c(p_lead1_sens, NA, p_lead1, NA, pp_lead1_sens, NA, pp_lead1, NA),3),
  RRR= c(sprintf("%.2f (%.2f–%.2f)\n", RRR_lead1_sens, lo95_lead1_sens, hi95_lead1_sens), NA, sprintf("%.2f (%.2f–%.2f)\n", RRR_lead1, lo95_lead1, hi95_lead1), NA, sprintf("%.2f (%.2f–%.2f)\n", RRR_p_lead1_sens, lo95_p_lead1_sens, hi95_p_lead1_sens), NA, sprintf("%.2f (%.2f–%.2f)\n", RRR_p_lead1, lo95_p_lead1, hi95_p_lead1), NA))|> 
  mutate(RR= sprintf("%.2f (%.2f–%.2f)", `exp(Est.)`, `2.5%`, `97.5%`))|>
  mutate(sig= sprintf("%.3f", sig))|> 
  rename("p.value"="P")|> 
  dplyr::select(term, p.value, RR, sig, RRR)|> 
  knitr::kable("markdown", caption="Main, high-speed collisions after one day", digits=3, na= "-")

cbind.data.frame(type=c("Main","","", "", "Placebo","","",""),
rbind.data.frame(
                tidy_exp_gnm_corr(model_gnm_tratados_lead2_sens),
                tidy_exp_gnm_corr(model_gnm_controles_lead2_sens),
                tidy_exp_gnm_corr(model_gnm_tratados_lead2),
                tidy_exp_gnm_corr(model_gnm_controles_lead2),
                tidy_exp_gnm_corr(placebo_model_gnm_tratados_lead2_sens),
                tidy_exp_gnm_corr(placebo_model_gnm_controles_lead2_sens),
                tidy_exp_gnm_corr(placebo_model_gnm_tratados_lead2),
                tidy_exp_gnm_corr(placebo_model_gnm_controles_lead2)),
  sig= round(c(p_lead2_sens, NA, p_lead2, NA, pp_lead2_sens, NA, pp_lead2, NA),3),
  RRR= c(sprintf("%.2f (%.2f–%.2f)", RRR_lead2_sens, lo95_lead2_sens, hi95_lead2_sens), NA, sprintf("%.2f (%.2f–%.2f)\n", RRR_lead2, lo95_lead2, hi95_lead2), NA, sprintf("%.2f (%.2f–%.2f)\n", RRR_p_lead2_sens, lo95_p_lead2_sens, hi95_p_lead2_sens), NA, sprintf("%.2f (%.2f–%.2f)\n", RRR_p_lead2, lo95_p_lead2, hi95_p_lead2), NA))|> 
  mutate(RR= sprintf("%.2f (%.2f–%.2f)", `exp(Est.)`, `2.5%`, `97.5%`))|>
  mutate(sig= sprintf("%.3f", sig))|> 
  rename("p.value"="P")|> 
  dplyr::select(term, p.value, RR, sig, RRR)|> 
  knitr::kable("markdown", caption="Main, high-speed collisions after two days", digits=3, na= "-")

pqic_gnm_tratados             <- qic_gnm(placebo_model_gnm_tratados)
pqic_gnm_tratados_lead1       <- qic_gnm(placebo_model_gnm_tratados_lead1)
pqic_gnm_tratados_lead2       <- qic_gnm(placebo_model_gnm_tratados_lead2)
pqic_gnm_tratados_sens        <- qic_gnm(placebo_model_gnm_tratados_sens)
pqic_gnm_tratados_lead1_sens  <- qic_gnm(placebo_model_gnm_tratados_lead1_sens)
pqic_gnm_tratados_lead2_sens  <- qic_gnm(placebo_model_gnm_tratados_lead2_sens)

pqic_gnm_controles            <- qic_gnm(placebo_model_gnm_controles)
pqic_gnm_controles_lead1      <- qic_gnm(placebo_model_gnm_controles_lead1)
pqic_gnm_controles_lead2      <- qic_gnm(placebo_model_gnm_controles_lead2)
pqic_gnm_controles_sens       <- qic_gnm(placebo_model_gnm_controles_sens)
pqic_gnm_controles_lead1_sens <- qic_gnm(placebo_model_gnm_controles_lead1_sens)
pqic_gnm_controles_lead2_sens <- qic_gnm(placebo_model_gnm_controles_lead2_sens)

## ─── 2. Vector with names ───────────────────────────────────────────
pqic_vec <- c(
  "Treated (base)"            = pqic_gnm_tratados,
  "Treated + lead 1"          = pqic_gnm_tratados_lead1,
  "Treated + lead 2"          = pqic_gnm_tratados_lead2,
  "Treated sensitivity (- Sherbrooke)"      = pqic_gnm_tratados_sens,
  "Treated sens + lead 1"     = pqic_gnm_tratados_lead1_sens,
  "Treated sens + lead 2"     = pqic_gnm_tratados_lead2_sens,
  "Controls (base)"           = pqic_gnm_controles,
  "Controls + lead 1"         = pqic_gnm_controles_lead1,
  "Controls + lead 2"         = pqic_gnm_controles_lead2,
  "Controls sensitivity (+ Sherbrooke)"     = pqic_gnm_controles_sens,
  "Controls sens + lead 1"    = pqic_gnm_controles_lead1_sens,
  "Controls sens + lead 2"    = pqic_gnm_controles_lead2_sens
)

## ─── 3. print results ───────────────────────────────────────
cat(
  "\n≡ Comparison of models through QIC (quasi‑AIC) ≡",
  "\n(‑2·ℓ / φ  +  2·k ;  lower ⇒ better)\n\n",
  paste(names(pqic_vec), sprintf(": %.1f", pqic_vec), collapse = "\n"),
  "\n\n≡ Ranking (better → worse) ≡\n",
  paste(names(sort(pqic_vec)), sprintf(": %.1f", sort(pqic_vec)), collapse = "\n"),
  "\n"
)

```


```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: test-association-gnm-main-vs-placebo
#| eval: true
#| results: "hold"


z_lead2_placebo_main_h   <- (b2_lead2 - pb2_lead2) / sqrt(se2_lead2^2 + pse2_lead2^2)
p_lead2_placebo_main_h   <- 2 * pnorm(-abs(z_lead2_placebo_main_h))     # 0.32
cat("Host MRC-years; Main vs. Placebo (w/o Sherbrooke), effects after 2 days;  p-values\n")
p_lead2_placebo_main_h

log_rrr_placebo_main_h  <- b2_lead2- pb2_lead2
se_rrr_placebo_main_h   <- sqrt(se2_lead2^2 + pse2_lead2^2)   # var(Δ) = var1 + var2

## RRR y IC 95 %
RRR_lead2_placebo_main_h      <- exp(log_rrr_placebo_main_h)
lo95_lead2_placebo_main_h     <- exp(log_rrr_placebo_main_h - 1.96 * se_rrr_placebo_main_h)
hi95_lead2_placebo_main_h     <- exp(log_rrr_placebo_main_h + 1.96 * se_rrr_placebo_main_h)

cat(sprintf("%.2f (%.2f–%.2f) p= %.3f", RRR_lead2_placebo_main_h, lo95_lead2_placebo_main_h, hi95_lead2_placebo_main_h, p_lead2_placebo_main_h))


z_lead2_placebo_main_nh   <- (b1_lead2 - pb1_lead2) / sqrt(se1_lead2^2 + pse1_lead2^2)
p_lead2_placebo_main_nh   <- 2 * pnorm(-abs(z_lead2_placebo_main_nh))     # 0.32
cat("Non-Host MRC-years; Main vs. Placebo (w/o Sherbrooke), effects after 2 days;  p-values\n")
p_lead2_placebo_main_nh

log_rrr_placebo_main_nh  <- b1_lead2- pb1_lead2
se_rrr_placebo_main_nh   <- sqrt(se1_lead2^2 + pse1_lead2^2)   # var(Δ) = var1 + var2

## RRR y IC 95 %
RRR_lead2_placebo_main_nh      <- exp(log_rrr_placebo_main_nh)
lo95_lead2_placebo_main_nh     <- exp(log_rrr_placebo_main_nh - 1.96 * se_rrr_placebo_main_nh)
hi95_lead2_placebo_main_nh     <- exp(log_rrr_placebo_main_nh + 1.96 * se_rrr_placebo_main_nh)

cat(sprintf("%.2f (%.2f–%.2f) p= %.3f", RRR_lead2_placebo_main_nh, lo95_lead2_placebo_main_nh, hi95_lead2_placebo_main_nh, p_lead2_placebo_main_nh))



z_lead2_placebo_main_h_sens   <- (b2_lead2_sens - pb2_lead2_sens) / sqrt(se2_lead2_sens^2 + pse2_lead2_sens^2)
p_lead2_placebo_main_h_sens   <- 2 * pnorm(-abs(z_lead2_placebo_main_h_sens))     # 0.32
cat("Host MRC-years; Main vs. Placebo (w/ Sherbrooke as control), effects after 2 days;  p-values\n")
p_lead2_placebo_main_h_sens

log_rrr_placebo_main_h_sens  <- b2_lead2_sens- pb2_lead2_sens
se_rrr_placebo_main_h_sens   <- sqrt(se2_lead2_sens^2 + pse2_lead2_sens^2)   # var(Δ) = var1 + var2

## RRR y IC 95 %
RRR_lead2_placebo_main_h_sens      <- exp(log_rrr_placebo_main_h_sens)
lo95_lead2_placebo_main_h_sens     <- exp(log_rrr_placebo_main_h_sens - 1.96 * se_rrr_placebo_main_h_sens)
hi95_lead2_placebo_main_h_sens     <- exp(log_rrr_placebo_main_h_sens + 1.96 * se_rrr_placebo_main_h_sens)

cat(sprintf("%.2f (%.2f–%.2f) p= %.3f", RRR_lead2_placebo_main_h_sens, lo95_lead2_placebo_main_h_sens, hi95_lead2_placebo_main_h_sens, p_lead2_placebo_main_h_sens))


z_lead2_placebo_main_nh_sens   <- (b1_lead2_sens - pb1_lead2_sens) / sqrt(se1_lead2_sens^2 + pse1_lead2_sens^2)
p_lead2_placebo_main_nh_sens   <- 2 * pnorm(-abs(z_lead2_placebo_main_nh_sens))     # 0.32
cat("Non-Host MRC-years; Main vs. Placebo (w/ Sherbrooke as control), effects after 2 days;  p-values\n")
p_lead2_placebo_main_nh_sens

log_rrr_placebo_main_nh_sens  <- b1_lead2_sens- pb1_lead2_sens
se_rrr_placebo_main_nh_sens   <- sqrt(se1_lead2_sens^2 + pse1_lead2_sens^2)   # var(Δ) = var1 + var2

## RRR y IC 95 %
RRR_lead2_placebo_main_nh_sens      <- exp(log_rrr_placebo_main_nh_sens)
lo95_lead2_placebo_main_nh_sens     <- exp(log_rrr_placebo_main_nh_sens - 1.96 * se_rrr_placebo_main_nh_sens)
hi95_lead2_placebo_main_nh_sens     <- exp(log_rrr_placebo_main_nh_sens + 1.96 * se_rrr_placebo_main_nh_sens)

cat(sprintf("%.2f (%.2f–%.2f) p= %.3f", RRR_lead2_placebo_main_nh_sens, lo95_lead2_placebo_main_nh_sens, hi95_lead2_placebo_main_nh_sens, p_lead2_placebo_main_nh_sens))
```

We made model with interaction terms.

```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: test-association-gnm-int-post-conv
#| eval: true
#| results: "hold"

cat("DID (w/ Sherbrooke), effects after 2 days\n")
model_gnm_lead2_did_sens <- gnm(
    sum_velocidad_lead2 ~ relevel(factor(exposure_window), ref="Pre-exposure")*treated_sens+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
        offset(off_lic),
    eliminate = cluster,                            # <── “conditional” part
    family    = quasipoisson(link = "log"),
    data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Exp",exposure_window))
)
summary(model_gnm_lead2_did_sens)
cat("DID (w/o Sherbrooke), effects after 2 days\n")
model_gnm_lead2_did <- gnm(
    sum_velocidad_lead2 ~ relevel(factor(exposure_window), ref="Pre-exposure")*treated+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
        offset(off_lic),
    eliminate = cluster,                            # <── “conditional” part
    family    = quasipoisson(link = "log"),
    data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Exp",exposure_window))
)
summary(model_gnm_lead2_did)


cat("PlaceboDID (w/ Sherbrooke), effects after 2 days\n")
placebo_model_gnm_lead2_did_sens <- gnm(
    sum_velocidad_lead2 ~ relevel(factor(exposure_window), ref="Pre-exposure")*treated_sens+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
        offset(off_lic),
    eliminate = cluster,                            # <── “conditional” part
    family    = quasipoisson(link = "log"),
    data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Post",exposure_window))
)
summary(placebo_model_gnm_lead2_did_sens)


cat("PlaceboDID (w/o Sherbrooke), effects after 2 days\n")
placebo_model_gnm_lead2_did <- gnm(
    sum_velocidad_lead2 ~ relevel(factor(exposure_window), ref="Pre-exposure")*treated+ mean_min_temp_mean_lin + mean_max_temp_mean_lin + median_total_precip_median_lin + median_lag_2_prec_median_imp+# spline cúbico (suaviza)
        offset(off_lic),
    eliminate = cluster,                            # <── “conditional” part
    family    = quasipoisson(link = "log"),
    data      = subset(collisions_weather_corr_rect_pre_post, grepl("^Pre|^Post",exposure_window) )
)
summary(placebo_model_gnm_lead2_did)


tidy_exp_gnm <- function(model, conf.level = 0.95, use_t = FALSE) {
  td <- broom::tidy(model)
  alpha <- 1 - conf.level
  crit <- if (use_t) {
    stats::qt(1 - alpha/2, df = model$df.residual)
  } else {
    stats::qnorm(1 - alpha/2)
  }
    td %>%
    mutate(
      conf.low_log  = estimate - crit * std.error,
      conf.high_log = estimate + crit * std.error,
      rate_ratio    = exp(estimate),
      rr_conf.low   = exp(conf.low_log),
      rr_conf.high  = exp(conf.high_log)
    ) %>%
    dplyr::select(term, rate_ratio, rr_conf.low, rr_conf.high, std.error, p.value) |> 
    data.frame() |> 
    dplyr::slice(7)
}

# tidy_exp_gnm2 <- function(model, conf.level = 0.95) {
#   # Extrae coeficientes y errores estándar
#   td <- broom::tidy(model)
#   
#   # Z crítico
#   alpha <- 1 - conf.level
#   z <- stats::qnorm(1 - alpha/2)
#   
#   # Cálculo de IC en escala log y exponenciación
#   td %>%
#     mutate(
#       conf.low_log  = estimate - z * std.error,
#       conf.high_log = estimate + z * std.error,
#       rate_ratio    = exp(estimate),
#       rr_conf.low   = exp(conf.low_log),
#       rr_conf.high  = exp(conf.high_log)
#     ) %>%
#     dplyr::select(term, rate_ratio, rr_conf.low, rr_conf.high, std.error, p.value) |> 
#     data.frame() |> 
#     dplyr::slice(7)
# }
cbind.data.frame(type=c("Main","","Placebo",""),
rbind.data.frame(as_tibble(Epi::ci.exp(model_gnm_lead2_did_sens, subset=":", pval=T), rownames = "term"),
                 as_tibble(Epi::ci.exp(model_gnm_lead2_did, subset=":", pval=T), rownames = "term"),
                as_tibble(Epi::ci.exp(placebo_model_gnm_lead2_did_sens, subset=":", pval=T), rownames = "term"),
                as_tibble(Epi::ci.exp(placebo_model_gnm_lead2_did, subset=":", pval=T), rownames = "term")))|> 
  mutate(RR= sprintf("%.2f (%.2f–%.2f)", `exp(Est.)`, `2.5%`, `97.5%`))|>
  knitr::kable("markdown", caption="Interaction models, RRRs, high-speed collisions after two days", digits=4)
```


```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: test-association-gnm-int
#| eval: false
#| results: "hold"
# Model: Main -2,+3 exposure, Quebec only as control
cat("Instantaneous effect, Main -2,+3 exposure, Quebec only as control\n")
model_gnm_did <- gnm(
  velocidad ~ D + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did)
aic_model_gnm_did <- qic_gnm(model_gnm_did)

# Model: Main 0,+3 exposure, Quebec only as control
cat("Instantaneous effect, Main 0,+3 exposure, Quebec only as control\n")
model_gnm_did_off <- gnm(
  velocidad ~ D_off + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did_off)
aic_model_gnm_did_off <- qic_gnm(model_gnm_did_off)

# Model: Main -2,+3 exposure, Quebec+Sherbrooke as control
cat("Instantaneous effect, Main -2,+3 exposure, Quebec+Sherbrooke as control\n")
model_gnm_did_sens <- gnm(
  velocidad ~ D_sens + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did_sens)
aic_model_gnm_did_sens <- qic_gnm(model_gnm_did_sens)

# Model: Main 0,+3 exposure, Quebec+Sherbrooke as control
cat("Instantaneous effect, Main 0,+3 exposure, Quebec+Sherbrooke as control\n")
model_gnm_did_sens_off <- gnm(
  velocidad ~ D_sens_off + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did_sens_off)
aic_model_gnm_did_sens_off <- qic_gnm(model_gnm_did_sens_off)

# Model: Main -2,+7 exposure, Quebec only as control
cat("Instantaneous effect, Main -2,+7 exposure, Quebec only as control\n")
model_gnm_did7 <- gnm(
  velocidad ~ D7 + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did7)
aic_model_gnm_did7 <- qic_gnm(model_gnm_did7)

# Model: Main 0,+7 exposure, Quebec only as control
cat("Instantaneous effect, Main 0,+7 exposure, Quebec only as control\n")
model_gnm_did7_off <- gnm(
  velocidad ~ D7_off + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did7_off)
aic_model_gnm_did7_off <- qic_gnm(model_gnm_did7_off)

# Model: Main -2,+7 exposure, Quebec+Sherbrooke as control
cat("Instantaneous effect, Main -2,+7 exposure, Quebec+Sherbrooke as control\n")
model_gnm_did7_sens <- gnm(
  velocidad ~ D7_sens + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did7_sens)
aic_model_gnm_did7_sens <- qic_gnm(model_gnm_did7_sens)

# Model: Main 0,+7 exposure, Quebec+Sherbrooke as control
cat("Instantaneous effect, Main 0,+7 exposure, Quebec+Sherbrooke as control\n")
model_gnm_did7_sens_off <- gnm(
  velocidad ~ D7_sens_off + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did7_sens_off)
aic_model_gnm_did7_sens_off <- qic_gnm(model_gnm_did7_sens_off)


#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_

# Model: Main -2,+3 exposure, Quebec only as control
cat("Lead 1, Main -2,+3 exposure, Quebec only as control, lead1\n")
model_gnm_did_lead1 <- gnm(
  velocidad_lead1_post_conv_imp ~ D + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did_lead1)
aic_model_gnm_did_lead1 <- qic_gnm(model_gnm_did_lead1)

# Model: Main 0,+3 exposure, Quebec only as control
cat("Lead 1, Main 0,+3 exposure, Quebec only as control, lead1\n")
model_gnm_did_off_lead1 <- gnm(
  velocidad_lead1_post_conv_imp ~ D_off + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did_off_lead1)
aic_model_gnm_did_off_lead1 <- qic_gnm(model_gnm_did_off_lead1)

# Model: Main -2,+3 exposure, Quebec+Sherbrooke as control
cat("Lead 1, Main -2,+3 exposure, Quebec+Sherbrooke as control, lead1\n")
model_gnm_did_sens_lead1 <- gnm(
  velocidad_lead1_post_conv_imp ~ D_sens + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did_sens_lead1)
aic_model_gnm_did_sens_lead1 <- qic_gnm(model_gnm_did_sens_lead1)

# Model: Main 0,+3 exposure, Quebec+Sherbrooke as control
cat("Lead 1, Main 0,+3 exposure, Quebec+Sherbrooke as control, lead1\n")
model_gnm_did_sens_off_lead1 <- gnm(
  velocidad_lead1_post_conv_imp ~ D_sens_off + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did_sens_off_lead1)
aic_model_gnm_did_sens_off_lead1 <- qic_gnm(model_gnm_did_sens_off_lead1)

# Model: Main -2,+7 exposure, Quebec only as control
cat("Lead 1, Main -2,+7 exposure, Quebec only as control, lead1\n")
model_gnm_did7_lead1 <- gnm(
  velocidad_lead1_post_conv_imp ~ D7 + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did7_lead1)
aic_model_gnm_did7_lead1 <- qic_gnm(model_gnm_did7_lead1)

# Model: Main 0,+7 exposure, Quebec only as control
cat("Lead 1, Main 0,+7 exposure, Quebec only as control, lead1\n")
model_gnm_did7_off_lead1 <- gnm(
  velocidad_lead1_post_conv_imp ~ D7_off + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did7_off_lead1)
aic_model_gnm_did7_off_lead1 <- qic_gnm(model_gnm_did7_off_lead1)

# Model: Main -2,+7 exposure, Quebec+Sherbrooke as control
cat("Lead 1, Main -2,+7 exposure, Quebec+Sherbrooke as control, lead1\n")
model_gnm_did7_sens_lead1 <- gnm(
  velocidad_lead1_post_conv_imp ~ D7_sens + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did7_sens_lead1)
aic_model_gnm_did7_sens_lead1 <- qic_gnm(model_gnm_did7_sens_lead1)

# Model: Main 0,+7 exposure, Quebec+Sherbrooke as control
cat("Lead 1, Main 0,+7 exposure, Quebec+Sherbrooke as control, lead1\n")
model_gnm_did7_sens_off_lead1 <- gnm(
  velocidad_lead1_post_conv_imp ~ D7_sens_off + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did7_sens_off_lead1)
aic_model_gnm_did7_sens_off_lead1 <- qic_gnm(model_gnm_did7_sens_off_lead1)



#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_
#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_#_

# Model: Main -2,+3 exposure, Quebec only as control
cat("Lead 2, Main -2,+3 exposure, Quebec only as control, lead2\n")
model_gnm_did_lead2 <- gnm(
  velocidad_lead2_post_conv_imp ~ D + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did_lead2)
aic_model_gnm_did_lead2 <- qic_gnm(model_gnm_did_lead2)

# Model: Main 0,+3 exposure, Quebec only as control
cat("Lead 2, Main 0,+3 exposure, Quebec only as control, lead2\n")
model_gnm_did_off_lead2 <- gnm(
  velocidad_lead2_post_conv_imp ~ D_off + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did_off_lead2)
aic_model_gnm_did_off_lead2 <- qic_gnm(model_gnm_did_off_lead2)

# Model: Main -2,+3 exposure, Quebec+Sherbrooke as control
cat("Lead 2, Main -2,+3 exposure, Quebec+Sherbrooke as control, lead2\n")
model_gnm_did_sens_lead2 <- gnm(
  velocidad_lead2_post_conv_imp ~ D_sens + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did_sens_lead2)
aic_model_gnm_did_sens_lead2 <- qic_gnm(model_gnm_did_sens_lead2)

# Model: Main 0,+3 exposure, Quebec+Sherbrooke as control
cat("Lead 2, Main 0,+3 exposure, Quebec+Sherbrooke as control, lead2\n")
model_gnm_did_sens_off_lead2 <- gnm(
  velocidad_lead2_post_conv_imp ~ D_sens_off + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did_sens_off_lead2)
aic_model_gnm_did_sens_off_lead2 <- qic_gnm(model_gnm_did_sens_off_lead2)

# Model: Main -2,+7 exposure, Quebec only as control
cat("Lead 2, Main -2,+7 exposure, Quebec only as control, lead2\n")
model_gnm_did7_lead2 <- gnm(
  velocidad_lead2_post_conv_imp ~ D7 + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did7_lead2)
aic_model_gnm_did7_lead2 <- qic_gnm(model_gnm_did7_lead2)

# Model: Main 0,+7 exposure, Quebec only as control
cat("Lead 2, Main 0,+7 exposure, Quebec only as control, lead2\n")
model_gnm_did7_off_lead2 <- gnm(
  velocidad_lead2_post_conv_imp ~ D7_off + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did7_off_lead2)
aic_model_gnm_did7_off_lead2 <- qic_gnm(model_gnm_did7_off_lead2)

# Model: Main -2,+7 exposure, Quebec+Sherbrooke as control
cat("Lead 2, Main -2,+7 exposure, Quebec+Sherbrooke as control, lead2\n")
model_gnm_did7_sens_lead2 <- gnm(
  velocidad_lead2_post_conv_imp ~ D7_sens + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did7_sens_lead2)
aic_model_gnm_did7_sens_lead2 <- qic_gnm(model_gnm_did7_sens_lead2)

# Model: Main 0,+7 exposure, Quebec+Sherbrooke as control
cat("Lead 2, Main 0,+7 exposure, Quebec+Sherbrooke as control, lead2\n")
model_gnm_did7_sens_off_lead2 <- gnm(
  velocidad_lead2_post_conv_imp ~ D7_sens_off + min_temp_mean_lin + max_temp_mean_lin + total_precip_median_lin + lag_2_prec_median_lin_imp + 
    offset(license_holders_sex_age),
  eliminate = cluster,
  family = quasipoisson(link = "log"),
  data = collisions_weather_corr
)
summary(model_gnm_did7_sens_off_lead2)
aic_model_gnm_did7_sens_off_lead2 <- qic_gnm(model_gnm_did7_sens_off_lead2)


# Print all AIC values at the end
cat("\nAIC Values for All Models:\n")
cat("Instantaneous effect, Main -2,+3 exposure, Quebec only as control:", aic_model_gnm_did, "\n")
cat("Instantaneous effect, Main 0,+3 exposure, Quebec only as control:", aic_model_gnm_did_off, "\n")
cat("Instantaneous effect, Main -2,+3 exposure, Quebec+Sherbrooke as control:", aic_model_gnm_did_sens, "\n")
cat("Instantaneous effect, Main 0,+3 exposure, Quebec+Sherbrooke as control:", aic_model_gnm_did_sens_off, "\n")
cat("Instantaneous effect, Main -2,+7 exposure, Quebec only as control:", aic_model_gnm_did7, "\n")
cat("Instantaneous effect, Main 0,+7 exposure, Quebec only as control:", aic_model_gnm_did7_off, "\n")
cat("Instantaneous effect, Main -2,+7 exposure, Quebec+Sherbrooke as control:", aic_model_gnm_did7_sens, "\n")
cat("Instantaneous effect, Main 0,+7 exposure, Quebec+Sherbrooke as control:", aic_model_gnm_did7_sens_off, "\n")

cat("Lead 1, Main -2,+3 exposure, Quebec only as control:", aic_model_gnm_did_lead1, "\n")
cat("Lead 1, Main 0,+3 exposure, Quebec only as control:", aic_model_gnm_did_off_lead1, "\n")
cat("Lead 1, Main -2,+3 exposure, Quebec+Sherbrooke as control:", aic_model_gnm_did_sens_lead1, "\n")
cat("Lead 1, Main 0,+3 exposure, Quebec+Sherbrooke as control:", aic_model_gnm_did_sens_off_lead1, "\n")
cat("Lead 1, Main -2,+7 exposure, Quebec only as control:", aic_model_gnm_did7_lead1, "\n")
cat("Lead 1, Main 0,+7 exposure, Quebec only as control:", aic_model_gnm_did7_off_lead1, "\n")
cat("Lead 1, Main -2,+7 exposure, Quebec+Sherbrooke as control:", aic_model_gnm_did7_sens_lead1, "\n")
cat("Lead 1, Main 0,+7 exposure, Quebec+Sherbrooke as control:", aic_model_gnm_did7_sens_off_lead1, "\n")

cat("Lead 2, Main -2,+3 exposure, Quebec only as control:", aic_model_gnm_did_lead2, "\n")
cat("Lead 2, Main 0,+3 exposure, Quebec only as control:", aic_model_gnm_did_off_lead2, "\n")
cat("Lead 2, Main -2,+3 exposure, Quebec+Sherbrooke as control:", aic_model_gnm_did_sens_lead2, "\n")
cat("Lead 2, Main 0,+3 exposure, Quebec+Sherbrooke as control:", aic_model_gnm_did_sens_off_lead2, "\n")
cat("Lead 2, Main -2,+7 exposure, Quebec only as control:", aic_model_gnm_did7_lead2, "\n")
cat("Lead 2, Main 0,+7 exposure, Quebec only as control:", aic_model_gnm_did7_off_lead2, "\n")
cat("Lead 2, Main -2,+7 exposure, Quebec+Sherbrooke as control:", aic_model_gnm_did7_sens_lead2, "\n")
cat("Lead 2, Main 0,+7 exposure, Quebec+Sherbrooke as control:", aic_model_gnm_did7_sens_off_lead2, "\n")
```



### Synth


```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: synth
#| eval: true
#| results: "hold"

cat("Parameters\n")
t_int_man<- abs(min_value)+1
max_time <- max(collisions_weather_corr_rect_synth$yday_corr) #38 #34 #number of days restricted


#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
cat("Print ID labs with treatment")

collisions_weather_corr_rect_synth_one_tr_many_cntrls$id_lab <- ifelse(is.na(collisions_weather_corr_rect_synth_one_tr_many_cntrls$id_lab), "treatment", collisions_weather_corr_rect_synth_one_tr_many_cntrls$id_lab)

collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls$id_lab <- ifelse(is.na(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls$id_lab), "treatment", collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls$id_lab)


#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:


cat("Added the Lag Y by status and year:\n")
collisions_weather_corr_rect_synth_one_tr_many_cntrls$Y_lag10 <- 
collisions_weather_corr_rect_synth_one_tr_many_cntrls|> 
    group_by(tr_contr_sens, year.x)|> 
    mutate(Y_lag10 = dplyr::lag(rate_veh, 10, order_by = yday_corr))|> 
    ungroup()|>
    pull(Y_lag10) 
collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls$Y_lag10 <- 
collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls|> 
    group_by(tr_contr, year.x)|> 
    mutate(Y_lag10 = dplyr::lag(rate_veh, 4, order_by = yday_corr))|> 
    ungroup()|>
    pull(Y_lag10) 
collisions_weather_corr_rect_synth_one_tr_many_cntrls$Y_lag10 <- 
collisions_weather_corr_rect_synth_one_tr_many_cntrls|>
  mutate(intersect= paste0(id,"_",year.x)) |> 
    group_by(intersect)|>
    mutate(
        third_value = nth(na.omit(Y_lag10), 1),
        Y_lag10 = ifelse(is.na(Y_lag10), third_value, Y_lag10)
    )|>
    dplyr::select(Y_lag10)|>
    ungroup()|>
    pull(Y_lag10)
collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls$Y_lag10 <- 
collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls|>
  mutate(intersect= paste0(id,"_",year.x)) |> 
    group_by(intersect)|>
    mutate(
        third_value = nth(na.omit(Y_lag10), 1),
        Y_lag10 = ifelse(is.na(Y_lag10), third_value, Y_lag10)
    )|>
    dplyr::select(Y_lag10)|>
    ungroup()|>
    pull(Y_lag10)


#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:
#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:#:

cat("*************************************\n")
cat("0, 38, Quebec+Sherbrooke as controls\n")
dataprep.out <- Synth::dataprep(
  foo = collisions_weather_corr_rect_synth_one_tr_many_cntrls |> data.frame(),
  predictors = c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", 
                 "mean_median_lag_2_prec_median_imp"),
  dependent = "rate_lead1_lic",
  unit.variable = "id",
  time.variable = "yday_corr",
  treatment.identifier = 46, # Laval, Montreal, Longueuil
  controls.identifier = 1:26,
  time.predictors.prior = 1:(t_int_man-1),
  time.optimize.ssr = 1:(t_int_man-1),
  time.plot= 1:max_time,#
  unit.names.variable = "id_lab"
)
synth.out <- synth(dataprep.out, method = "BFGS")
print(synth.out)

#discrepancies between control and synthetic control
gaps <- dataprep.out$Y1plot - (dataprep.out$Y0plot %*% synth.out$solution.w)
synth.tables <- synth.tab(dataprep.res = dataprep.out,
                       synth.res    = synth.out
                       )

cat("*************************************\n")
cat("0, 38, Quebec only as controls\n")
cat("For quebec only as a control\n")
dataprep.out_q <- Synth::dataprep(
  foo = collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls |> data.frame(),
  predictors = c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", 
                 "mean_median_lag_2_prec_median_imp"),
  dependent = "rate_veh",
  unit.variable = "id",
  time.variable = "yday_corr",
  treatment.identifier = 46, # Laval, Montreal, Longueuil
  controls.identifier = 1:26,
  time.predictors.prior = 1:(t_int_man-1),
  time.optimize.ssr = 1:(t_int_man-1),
  time.plot= 1:max_time,
  unit.names.variable = "id_lab"
)
synth.out_q <- synth(dataprep.out_q, method = "BFGS")
print(synth.out_q)

#discrepancies between control and synthetic control
gaps_q <- dataprep.out_q$Y1plot - (dataprep.out_q$Y0plot %*% synth.out_q$solution.w)
synth.tables_q <- synth.tab(dataprep.res = dataprep.out_q,
                       synth.res    = synth.out_q
                       )

rbind.data.frame(cbind.data.frame(type="0-38, Quebec+Sherbrooke as controls",synth.tables$tab.pred,  predictor_weight=synth.tables$tab.v), cbind.data.frame(type="0-38, Quebec only as control",synth.tables_q$tab.pred,  predictor_weight=synth.tables_q$tab.v))|> 
  mutate(`Sample Mean`= sprintf("%1.2f",`Sample Mean`), v.weights= sprintf("%1.2f",v.weights), abs_diff= sprintf("%1.2f",abs(Treated-Synthetic)), ratio= sprintf("%1.2f",Treated/Synthetic), gap_pct_ctrl= sprintf("%1.2f", abs(100*(Treated-Synthetic)/Synthetic)))|> 
  knitr::kable("markdown", caption= "Comparison of pre-treatment predictor values for treated unit, synthetic control and all units in the sample")
rbind.data.frame(cbind.data.frame(type="0-38, Quebec+Sherbrooke as controls", synth.tables$tab.w), cbind.data.frame(type="0-38, Quebec only as control", synth.tables_q$tab.w))|> 
  knitr::kable("markdown", caption="Weight for each potential control unit")

path.plot(synth.res = synth.out_q,
          dataprep.res = dataprep.out_q,
          Main= "0-38, Quebec only as control",
          Ylab = "Rate per Vehicle",
          Xlab = "Time Period",
          Ylim = c(0,3),
          Legend = c("Treated cities","synthetic treated cities"),
          Legend.position = "topright",
          tr.intake = 38
)
gaps.plot(synth.res = synth.out_q, dataprep.res = dataprep.out_q,
          Ylab = "Gap in Rate", Xlab = "Time Period",
          Main = "Treatment Effect (difference over time)\n[0-38, Quebec only as control]",
          tr.intake = 38)

path.plot(synth.res = synth.out,
          dataprep.res = dataprep.out,
          Main= "0-38, Quebec+Sherbrooke as controls",
          Ylab = "Rate per Vehicle",
          Xlab = "Time Period",
          Ylim = c(0,3),
          Legend = c("Treated cities","synthetic treated cities"),
          Legend.position = "topright",
          tr.intake = 38
)
gaps.plot(synth.res = synth.out, dataprep.res = dataprep.out,
          Ylab = "Gap in Rate", Xlab = "Time Period",
          Main = "Treatment Effect (difference over time)\n[0-38, Sherbrooke+Quebec as controls]",
          tr.intake = 38)

cat("**********************************************\n")
cat("Placebo tests\n")

tdf <- SCtools::generate.placebos(dataprep.out_q,synth.out_q, Sigf.ipop = 2) #The Precision setting for the ipop optimization routine. Default of 5.
tdf_sens <- SCtools::generate.placebos(dataprep.out,synth.out, Sigf.ipop = 2) #The Precision setting for the ipop optimization routine. Default of 5.

SCtools::plot_placebos(tdf_sens,title= "0-38, Quebec only as control")

SCtools::mspe.test(tdf_sens)$p.val#proportion of ratios of RMSPE, <.05 is significant

SCtools::mspe.plot(tdf = tdf_sens, title= "0-38, Quebec only as control\nRatios of RMSPE")


SCtools::plot_placebos(tdf,title= "0-38, Sherbrooke+Quebec as controls")

SCtools::mspe.test(tdf)$p.val #proportion of ratios of RMSPE, <.05 is significant

SCtools::mspe.plot(tdf = tdf, title= "0-38, Sherbrooke+Quebec as controls\nRatios of RMSPE")
```

- The pre-treatment balance diagnostics show an excellent fit between treated and its synthetic control.  The absolute gaps in the main predictors translate into relative differences of ≤ 10 %.  Crucially, the two variables that receive virtually all the weight in the V-matrix—current precipitation and precipitation lagged two days—match almost perfectly (≤ 0.3 % gap), as does the 10-day lag of the outcome. Overall, the synthetic control replicates the treated unit’s pre-grand prix trajectory very closely, so the absence of a post-event divergence cannot be attributed to poor pre-treatment balance.*
- Units 6, 7, 14, 20 explain more than 97% of the synthetic control (w.weights). For Quebec only as controls, donor units 7, 15, 6, 20, 2 explain 79% of the synthetic control.
- The effect post Grand Prix does not differentiate compared to the pseudo-treated units. There is no evidence of an unusual causal effect.
- Ratios of RMSPE (root mean squared prediction errors) post/pre-intervention and the proportion of control units with ratios greater or equal to treated does not detect a causal effect.

We replicated the analysis with the intervention period **2 days before the race**.


```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: synth2
#| eval: true
#| results: hold
if(
identical(collisions_weather_corr$total_precip_median_lin, collisions_weather_corr$lag_2_prec_median_lin)
){stop("The variable was not lagged")}

collisions_weather_corr_rect_synth_one_tr_many_cntrls |> 
    filter(is.na(Y_lag10)) |> nrow()


collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls |> 
    filter(is.na(Y_lag10)) |> nrow()

collisions_weather_corr_rect_synth_one_tr_many_cntrls$id <- as.numeric(collisions_weather_corr_rect_synth_one_tr_many_cntrls$id)
collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls$id <- as.numeric(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls$id)

if(max(collisions_weather_corr_rect_synth_one_tr_many_cntrls$rate_veh)<1){
collisions_weather_corr_rect_synth_one_tr_many_cntrls$rate_veh <- collisions_weather_corr_rect_synth_one_tr_many_cntrls$rate_veh*10
}

if(max(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls$rate_veh)<1){
collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls$rate_veh <- collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls$rate_veh*10
}

collisions_weather_corr_rect_synth_one_tr_many_cntrls<- 
  data.frame(collisions_weather_corr_rect_synth_one_tr_many_cntrls)

collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls<- data.frame(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls)


##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:
##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:
cat("*************************************\n")
cat("0, 35, Quebec+Sherbrooke as controls\n")
dataprep.out2 <- Synth::dataprep(
  foo = collisions_weather_corr_rect_synth_one_tr_many_cntrls,
  predictors = c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", 
                 "mean_median_lag_2_prec_median_imp"),
  dependent = "rate_veh",
  unit.variable = "id",
  time.variable = "yday_corr",
  treatment.identifier = 46, # Laval, Montreal, Longueuil
  controls.identifier = 1:26,
  time.predictors.prior = 1:(t_int_man-3), #nates tenía menos 1
  time.optimize.ssr = 1:(t_int_man-3),
  time.plot= 1:max_time,
  unit.names.variable = "id_lab"
)
synth.out2 <- synth(dataprep.out2, method = "BFGS")
print(synth.out2)

gaps2 <- dataprep.out2$Y1plot - (dataprep.out2$Y0plot %*% synth.out2$solution.w)
synth.tables2 <- synth.tab(dataprep.res = dataprep.out2,
                       synth.res    = synth.out2
                       )

cat("*************************************\n")
cat("0, 35, Quebec only as control\n")
dataprep.out2_q <- Synth::dataprep(
  foo = collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls,
  predictors = c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", 
                 "mean_median_lag_2_prec_median_imp"),
  dependent = "rate_veh",
  unit.variable = "id",
  time.variable = "yday_corr",
  treatment.identifier = 46, # Laval, Montreal, Longueuil
  controls.identifier = 1:26,
  time.predictors.prior = 1:(t_int_man-3),
  time.optimize.ssr = 1:(t_int_man-3),
  time.plot= 1:max_time,
  unit.names.variable = "id_lab"
)
synth.out2_q <- synth(dataprep.out2_q, method = "BFGS")
print(synth.out2_q)

gaps2_q <- dataprep.out2_q$Y1plot - (dataprep.out2_q$Y0plot %*% synth.out2_q$solution.w)
synth.tables2_q <- synth.tab(dataprep.res = dataprep.out2_q,
                       synth.res    = synth.out2_q
                       )


rbind.data.frame(cbind.data.frame(type="0-35, Quebec+Sherbrooke as controls",synth.tables2$tab.pred,  predictor_weight=synth.tables2$tab.v), cbind.data.frame(type="0-35, Quebec only as control",synth.tables2_q$tab.pred,  predictor_weight=synth.tables2_q$tab.v))|> 
  mutate(`Sample Mean`= sprintf("%1.2f",`Sample Mean`), v.weights= sprintf("%1.2f",v.weights), abs_diff= sprintf("%1.2f",abs(Treated-Synthetic)), ratio= sprintf("%1.2f",Treated/Synthetic), gap_pct_ctrl= sprintf("%1.2f", abs(100*(Treated-Synthetic)/Synthetic)))|> 
  knitr::kable("markdown", caption= "Comparison of pre-treatment predictor values for treated unit, synthetic control and all units in the sample")

rbind.data.frame(cbind.data.frame(type="0-35, Quebec+Sherbrooke as controls", synth.tables2$tab.w), cbind.data.frame(type="0-35, Quebec only as control", synth.tables2_q$tab.w))|>
  arrange(type, desc(w.weights))|>
  knitr::kable("markdown", caption="Weight for each potential control unit")

path.plot(synth.res = synth.out2_q,
          dataprep.res = dataprep.out2_q,
          Main= "0-35, Quebec only as control",
          Ylab = "Rate per Vehicle",
          Xlab = "Time Period",
          Ylim = c(0,3),
          Legend = c("Treated cities","synthetic treated cities"),
          Legend.position = "topright",
          tr.intake = 38-3
)
gaps.plot(synth.res = synth.out2_q, dataprep.res = dataprep.out2_q,
          Ylab = "Gap in Rate", Xlab = "Time Period",
          Main = "Treatment Effect (difference over time)\n[0-35, Quebec only as control]",
          tr.intake = 38-3)


path.plot(synth.res = synth.out2,
          dataprep.res = dataprep.out2,
          Main= "0-35, Quebec+Sherbrooke as controls",
          Ylab = "Rate per Vehicle",
          Xlab = "Time Period",
          Ylim = c(0,3),
          Legend = c("Treated cities","synthetic treated cities"),
          Legend.position = "topright",
          tr.intake = 38-3
)
gaps.plot(synth.res = synth.out2, dataprep.res = dataprep.out2,
          Ylab = "Gap in Rate", Xlab = "Time Period",
          Main = "Treatment Effect (difference over time)\n[0-35, Sherbrooke+Quebec as controls]",
          tr.intake = 38-3)

cat("**********************************************\n")
cat("Placebo tests\n")
tdf2_q <- SCtools::generate.placebos(dataprep.out2_q,synth.out2_q, Sigf.ipop = 2) #The Precision setting for the ipop optimization routine. Default of 5.
SCtools::plot_placebos(tdf2_q,title= "0-35, Quebec only as control")
SCtools::mspe.test(tdf2_q)$p.val
SCtools::mspe.plot(tdf = tdf2_q, title= "0-35, Quebec only as control\nRatios of RMSPE")


tdf2 <- SCtools::generate.placebos(dataprep.out2,synth.out2, Sigf.ipop = 2) #The Precision setting for the ipop optimization routine. Default of 5.
SCtools::plot_placebos(tdf2,title= "0-35, Sherbrooke+Quebec as controls")
SCtools::mspe.test(tdf2)$p.val
SCtools::mspe.plot(tdf = tdf2, title= "0-35, Sherbrooke+Quebec as controls\nRatios of RMSPE")
```

### CountSynth

We used a new **synthetic-control tool for counts and other non-negative numbers**. It matches the pre-treatment path like usual, but its error rule lets it stretch—while penalizing—any needed extrapolation. No log tricks or “+1” fixes are needed, so predictions stay non-negative even when the data include zeros  [@Bonander2021] [https://youtu.be/e5hmK5GzCHc](https://youtu.be/e5hmK5GzCHc).


#### Instantaneous effect

```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: synthcount
#| eval: true
#| results: hold
source(paste0(getwd(),"/_funcs/CSCM_helper_functions.R"))
#Bonander, Carl. A (Flexible) Synthetic Control Method for Count Data and Other Nonnegative Outcomes. Epidemiology 32(5):p 653-660, September 2021. | DOI: 10.1097/EDE.0000000000001388

packages <- c("ggplot2", "Synth", "glmnet",
              "dplyr", "osqp", "optimx")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}
sapply(packages, require, character.only=TRUE)
##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:
##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:##:#:#:#::#:#:#:#:

nrow(subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, id=="40", rate_veh))
set.seed(2125)
road.main.res <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_veh",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            unit.names= "year.x",
                            K=3) #Number of holdout periods for cross-fitting (try changing to 3 for sensitivity analysis) #Splits the pre-period into K blocks //Uses K-1 blocks for training when evaluating each holdout block // Ensures at least one time point remains for predictor calculation

set.seed(2125)
road.main.res_35 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_veh",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_veh",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_35 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_veh",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 


set.seed(2125)
road.main.res_38 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=(max_time-4)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_veh",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) #Number of holdout periods for cross-fitting (try changing to 3 for sensitivity analysis) #Splits the pre-period into K blocks //Uses K-1 blocks for training when evaluating each holdout block // Ensures at least one time point remains for predictor calculation

set.seed(2125)
road.main.res_35_38 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=(max_time-7)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_veh",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_38 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=(max_time-4)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_veh",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_35_38 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=(max_time-7)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_veh",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 


set.seed(2125)
road.main.res1se <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_veh",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) #Number of holdout periods for cross-fitting (try changing to 3 for sensitivity analysis) #Splits the pre-period into K blocks //Uses K-1 blocks for training when evaluating each holdout block // Ensures at least one time point remains for predictor calculation

set.seed(2125)
road.main.res_351se <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_veh",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q1se <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_veh",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_351se <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_veh",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 


set.seed(2125)
road.main.res_381se <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=(max_time-4)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_veh",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) #Number of holdout periods for cross-fitting (try changing to 3 for sensitivity analysis) #Splits the pre-period into K blocks //Uses K-1 blocks for training when evaluating each holdout block // Ensures at least one time point remains for predictor calculation

set.seed(2125)
road.main.res_35_381se <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=(max_time-7)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_veh",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_381se <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=(max_time-4)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_veh",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_35_381se <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=(max_time-7)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_veh",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 


cat("0,+7 ; Quebec+Sherbrook as control\n")
road.main.res$ATT
cat("-2,+7 ; Quebec+Sherbrook as control\n")
road.main.res_35$ATT
cat("0,+7 ; Quebec only as control\n")
road.main.res_q$ATT
cat("-2,+7 ; Quebec+Sherbrook as control\n")
road.main.res_35$ATT
cat("-3,+0 ; Quebec+Sherbrook as control\n")
road.main.res_35_38$ATT
cat("0,+3 ; Quebec only as control\n")
road.main.res_q_38$ATT
cat("-3,+0 ; Quebec only as control\n")
road.main.res_q_35_38$ATT

cat("Penalized, 0,+7 ; Quebec+Sherbrook as control\n")
road.main.res1se$ATT
cat("Penalized, -2,+7 ; Quebec+Sherbrook as control\n")
road.main.res_351se$ATT
cat("Penalized, 0,+7 ; Quebec only as control\n")
road.main.res_q1se$ATT
cat("Penalized, -2,+7 ; Quebec+Sherbrook as control\n")
road.main.res_351se$ATT
cat("Penalized, -3,+0 ; Quebec+Sherbrook as control\n")
road.main.res_35_381se$ATT
cat("Penalized, 0,+3 ; Quebec only as control\n")
road.main.res_q_381se$ATT
cat("Penalized, -3,+0 ; Quebec only as control\n")
road.main.res_q_35_381se$ATT
```


```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: synth3
#| eval: true
#| results: hold
#| fig.width: 12
#K= Number of non-overlapping hold-out blocks (“folds”) used for cross-fitting the weights and penalty. More folds ⇒ less bias but higher variance and longer run time.
#When countSynth tunes its ridge/elastic-net penalty (λ) by cross-validation, this flag tells it not to use the “one–standard-error” rule. If you set it to TRUE, the algorithm would pick the most parsimonious λ within 1 s.e. of the CV minimum.

#prepost_rmspe_cntsh(get_covbal_csnth(road.main.res), t_int=38)

result_objects <- list(
    list(obj = road.main.res, name = "road.main.res", desc = "0,+7 ; Quebec+Sherbrook as control"),
    list(obj = road.main.res_35, name = "road.main.res_35", desc = "-2,+7 ; Quebec+Sherbrook as control"),
    list(obj = road.main.res_q, name = "road.main.res_q", desc = "0,+7 ; Quebec only as control"),
    list(obj = road.main.res_q_35, name = "road.main.res_q_35", desc = "-2,+7 ; Quebec only as control (exRepeat)"),
    list(obj = road.main.res_35_38, name = "road.main.res_35_38", desc = "-3,+0 ; Quebec+Sherbrook as control"),
    list(obj = road.main.res_q_38, name = "road.main.res_q_38", desc = "0,+3 ; Quebec only as control"),
    list(obj = road.main.res_q_35_38, name = "road.main.res_q_35_38", desc = "-3,+0 ; Quebec only as control"),
    
    # Penalized Models
    list(obj = road.main.res1se, name = "road.main.res1se", desc = "Penalized, 0,+7 ; Quebec+Sherbrook as control"),
    list(obj = road.main.res_351se, name = "road.main.res_351se", desc = "Penalized, -2,+7 ; Quebec+Sherbrook as control"),
    list(obj = road.main.res_q1se, name = "road.main.res_q1se", desc = "Penalized, 0,+7 ; Quebec only as control"),
    list(obj = road.main.res_q_351se, name = "road.main.res_q_351se", desc = "Penalized, -2,+7 ; Quebec only as control (exRepeat)"),
    list(obj = road.main.res_35_381se, name = "road.main.res_35_381se", desc = "Penalized, -3,+0 ; Quebec+Sherbrook as control"),
    list(obj = road.main.res_q_381se, name = "road.main.res_q_381se", desc = "Penalized, 0,+3 ; Quebec only as control"),
    list(obj = road.main.res_q_35_381se, name = "road.main.res_q_35_381se", desc = "Penalized, -3,+0 ; Quebec only as control")
)


# --------------------------------------------------------------------------
# 3. Loop through objects, print ATT, and create/store plots
# --------------------------------------------------------------------------

# Initialize an empty list to store the generated ggplot objects.
plot_list <- list()

# This loop iterates over each item in the `result_objects` list.
for (item in result_objects) {

  # --- Print the ATT Value ---
  cat("--------------------------------------------------\n")
  cat("Processing:", item$name, "\n")
  cat(item$desc, "\n")
  print(item$obj$ATT)
  cat("\n")

  # --- Generate and Store the Plot ---
  # Check if the object exists to prevent errors
  if (exists(item$name) && !is.null(get(item$name))) {

    # Call the plotting function which returns a ggplot object
    figcscm <- plot_synth_cscm(
      res_object = item$obj,
      t_int_man = 38,       # Manual intervention time
      max_time = 45,        # Time horizon for the plot
      y_lab = "Crash rate", # Y-axis label
      colors = c("SCM" = "gray60", "CSCM" = "red", "Obs" = "black")
    )

    # Add the title using ggplot2's labs() function and store it
    plot_list[[item$name]] <- figcscm + labs(title = item$desc) + theme(plot.title = element_text(size = 13))

  } else {
    cat("Warning: Object", item$name, "not found. Skipping plot generation.\n")
  }
}

# --------------------------------------------------------------------------
# 4. Arrange and display the plots in 2x2 grids using cowplot
# --------------------------------------------------------------------------

# Now, print the combined plots. In a Quarto/Rmd file, this will render
# the grids directly in the output document.
print(plot_grid(plotlist = plot_list[1:4], ncol = 2, nrow = 2))

print(plot_grid(plotlist = plot_list[5:8], ncol = 2, nrow = 2))

print(plot_grid(plotlist = plot_list[9:12], ncol = 2, nrow = 2))

print(plot_grid(plotlist = plot_list[13:14], ncol = 2, nrow = 2))

cbind.data.frame(
    pen= c(rep(c("No","Yes"),each=2)),
    mods= c(rep(c("Quebec+Sherbrooke","Quebec only"),2)),
    rbind.data.frame(result_objects[[5]]$obj$ATT,result_objects[[7]]$obj$ATT, result_objects[[12]]$obj$ATT, result_objects[[14]]$obj$ATT)
) |> knitr::kable("markdown", caption="RRs")
```


#### Lagged effect (1 day)

```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: synthcount-lead1
#| eval: true
#| results: hold

set.seed(2125)
road.main.res_lead1 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead1_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            unit.names= "year.x",
                            K=3) #Number of holdout periods for cross-fitting (try changing to 3 for sensitivity analysis) #Splits the pre-period into K blocks //Uses K-1 blocks for training when evaluating each holdout block // Ensures at least one time point remains for predictor calculation

set.seed(2125)
road.main.res_35_lead1 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead1_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_lead1 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead1_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_35_lead1 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead1_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 


set.seed(2125)
road.main.res_38_lead1 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=(max_time-4)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead1_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) #Number of holdout periods for cross-fitting (try changing to 3 for sensitivity analysis) #Splits the pre-period into K blocks //Uses K-1 blocks for training when evaluating each holdout block // Ensures at least one time point remains for predictor calculation

set.seed(2125)
road.main.res_35_38_lead1 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=(max_time-7)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead1_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_38_lead1 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=(max_time-4)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead1_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_35_38_lead1 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=(max_time-7)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead1_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 


set.seed(2125)
road.main.res1se_lead1 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead1_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) #Number of holdout periods for cross-fitting (try changing to 3 for sensitivity analysis) #Splits the pre-period into K blocks //Uses K-1 blocks for training when evaluating each holdout block // Ensures at least one time point remains for predictor calculation

set.seed(2125)
road.main.res_351se_lead1 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead1_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q1se_lead1 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead1_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_351se_lead1 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead1_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 


set.seed(2125)
road.main.res_381se_lead1 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=(max_time-4)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead1_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) #Number of holdout periods for cross-fitting (try changing to 3 for sensitivity analysis) #Splits the pre-period into K blocks //Uses K-1 blocks for training when evaluating each holdout block // Ensures at least one time point remains for predictor calculation

set.seed(2125)
road.main.res_35_381se_lead1 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=(max_time-7)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead1_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_381se_lead1 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=(max_time-4)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead1_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_35_381se_lead1 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=(max_time-7)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead1_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

```



```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: synth3_lead1
#| eval: true
#| results: hold
#| fig.width: 12
#K= Number of non-overlapping hold-out blocks (“folds”) used for cross-fitting the weights and penalty. More folds ⇒ less bias but higher variance and longer run time.
#When countSynth tunes its ridge/elastic-net penalty (λ) by cross-validation, this flag tells it not to use the “one–standard-error” rule. If you set it to TRUE, the algorithm would pick the most parsimonious λ within 1 s.e. of the CV minimum.

#prepost_rmspe_cntsh(get_covbal_csnth(road.main.res), t_int=38)

result_objects2 <- list(
  list(obj = road.main.res_lead1, name = "road.main.res_lead1", desc = "0,+7 ; Quebec+Sherbrook as control"),
  list(obj = road.main.res_35_lead1, name = "road.main.res_35_lead1", desc = "-2,+7 ; Quebec+Sherbrook as control"),
  list(obj = road.main.res_q_lead1, name = "road.main.res_q_lead1", desc = "0,+7 ; Quebec only as control"),
  list(obj = road.main.res_q_35_lead1, name = "road.main.res_q_35_lead1", desc = "-2,+7 ; Quebec only as control (exRepeat)"),
  list(obj = road.main.res_35_38_lead1, name = "road.main.res_35_38_lead1", desc = "-3,+0 ; Quebec+Sherbrook as control"),
  list(obj = road.main.res_q_38_lead1, name = "road.main.res_q_38_lead1", desc = "0,+3 ; Quebec only as control"),
  list(obj = road.main.res_q_35_38_lead1, name = "road.main.res_q_35_38_lead1", desc = "-3,+0 ; Quebec only as control"),

  # Penalized Models
  list(obj = road.main.res1se_lead1, name = "road.main.res1se_lead1", desc = "Penalized, 0,+7 ; Quebec+Sherbrook as control"),
  list(obj = road.main.res_351se_lead1, name = "road.main.res_351se_lead1", desc = "Penalized, -2,+7 ; Quebec+Sherbrook as control"),
  list(obj = road.main.res_q1se_lead1, name = "road.main.res_q1se_lead1", desc = "Penalized, 0,+7 ; Quebec only as control"),
  list(obj = road.main.res_q_351se_lead1, name = "road.main.res_q_351se_lead1", desc = "Penalized, -2,+7 ; Quebec only as control (exRepeat)"),
  list(obj = road.main.res_35_381se_lead1, name = "road.main.res_35_381se_lead1", desc = "Penalized, -3,+0 ; Quebec+Sherbrook as control"),
  list(obj = road.main.res_q_381se_lead1, name = "road.main.res_q_381se_lead1", desc = "Penalized, 0,+3 ; Quebec only as control"),
  list(obj = road.main.res_q_35_381se_lead1, name = "road.main.res_q_35_381se_lead1", desc = "Penalized, -3,+0 ; Quebec only as control")
)

# --------------------------------------------------------------------------
# 3. Loop through objects, print ATT, and create/store plots
# --------------------------------------------------------------------------

# Initialize an empty list to store the generated ggplot objects.
plot_list2 <- list()

# This loop iterates over each item in the `result_objects` list.
for (item in result_objects2) {

  # --- Print the ATT Value ---
  cat("--------------------------------------------------\n")
  cat("Processing:", item$name, "\n")
  cat(item$desc, "\n")
  print(item$obj$ATT)
  cat("\n")

  # --- Generate and Store the Plot ---
  # Check if the object exists to prevent errors
  if (exists(item$name) && !is.null(get(item$name))) {

    # Call the plotting function which returns a ggplot object
    figcscm <- plot_synth_cscm(
      res_object = item$obj,
      t_int_man = 38,       # Manual intervention time
      max_time = 45,        # Time horizon for the plot
      y_lab = "Crash rate", # Y-axis label
      colors = c("SCM" = "gray60", "CSCM" = "red", "Obs" = "black")
    )

    # Add the title using ggplot2's labs() function and store it
    plot_list2[[item$name]] <- figcscm + labs(title = item$desc) + theme(plot.title = element_text(size = 13))

  } else {
    cat("Warning: Object", item$name, "not found. Skipping plot generation.\n")
  }
}

# --------------------------------------------------------------------------
# 4. Arrange and display the plots in 2x2 grids using cowplot
# --------------------------------------------------------------------------

# Now, print the combined plots. In a Quarto/Rmd file, this will render
# the grids directly in the output document.
print(plot_grid(plotlist = plot_list2[1:4], ncol = 2, nrow = 2))

print(plot_grid(plotlist = plot_list2[5:8], ncol = 2, nrow = 2))

print(plot_grid(plotlist = plot_list2[9:12], ncol = 2, nrow = 2))

print(plot_grid(plotlist = plot_list2[13:14], ncol = 2, nrow = 2))

cbind.data.frame(
    pen= c(rep(c("No","Yes"),each=2)),
    mods= c(rep(c("Quebec+Sherbrooke","Quebec only"),2)),
    rbind.data.frame(result_objects2[[5]]$obj$ATT,result_objects2[[7]]$obj$ATT, result_objects2[[12]]$obj$ATT, result_objects2[[14]]$obj$ATT)
) |> knitr::kable("markdown", caption="RRs")
```


#### Lagged effect (2 days)

```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: synthcount-lead2
#| eval: true
#| results: hold

set.seed(2125)
road.main.res_lead2 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead2_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            unit.names= "year.x",
                            K=3) #Number of holdout periods for cross-fitting (try changing to 3 for sensitivity analysis) #Splits the pre-period into K blocks //Uses K-1 blocks for training when evaluating each holdout block // Ensures at least one time point remains for predictor calculation

set.seed(2125)
road.main.res_35_lead2 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead2_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_lead2 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead2_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_35_lead2 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead2_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 


set.seed(2125)
road.main.res_38_lead2 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead2_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) #Number of holdout periods for cross-fitting (try changing to 3 for sensitivity analysis) #Splits the pre-period into K blocks //Uses K-1 blocks for training when evaluating each holdout block // Ensures at least one time point remains for predictor calculation

set.seed(2125)
road.main.res_35_38_lead2 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=(max_time-7)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead2_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_38_lead2 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead2_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_35_38_lead2 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=(max_time-7)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead2_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= F, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 


set.seed(2125)
road.main.res1se_lead2 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead2_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) #Number of holdout periods for cross-fitting (try changing to 3 for sensitivity analysis) #Splits the pre-period into K blocks //Uses K-1 blocks for training when evaluating each holdout block // Ensures at least one time point remains for predictor calculation

set.seed(2125)
road.main.res_351se_lead2 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead2_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q1se_lead2 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead2_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_351se_lead2 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead2_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 


set.seed(2125)
road.main.res_381se_lead2 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=(max_time-7)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead2_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) #Number of holdout periods for cross-fitting (try changing to 3 for sensitivity analysis) #Splits the pre-period into K blocks //Uses K-1 blocks for training when evaluating each holdout block // Ensures at least one time point remains for predictor calculation

set.seed(2125)
road.main.res_35_381se_lead2 <- countSynth(data= subset(collisions_weather_corr_rect_synth_one_tr_many_cntrls, yday_corr<=(max_time-7)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead2_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_381se_lead2 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=max_time),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead2_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 38, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 

set.seed(2125)
road.main.res_q_35_381se_lead2 <- countSynth(data= subset(collisions_weather_corr_rect_quebec_synth_one_tr_many_cntrls, yday_corr<=(max_time-7)),
                            predictors=c("mean_min_temp_mean_lin", "mean_max_temp_mean_lin", "mean_median_total_precip_median_lin", "mean_median_lag_2_prec_median_imp"), #Auxiliary covariates can be included here as well (passed to dataprep synth, averaged across pre-period with na.rm=TRUE). Caution: needs to have non-missing values in all holdout samples.
                            dependent="rate_lead2_lic",
                            unit.variable="id",
                            time.variable = "yday_corr",
                            treatment.identifier = 46, # 25 = Sweden
                            controls.identifier = 1:26, #We exclude UK, Norway, Netherlands due to similar policies. Est, Rou, Lva due to data issues. Other countries due to missing values.
                            t_int= 35, #Treatment time point (1997 in real years)
                            min_1se= T, #Use 1se rule for lambda CV? (try changing to T for sensitivity analysis)
                            full.model = TRUE,
                            K=4) 


cat("0,+7 ; Quebec+Sherbrook as control\n")
road.main.res_lead2$ATT
cat("-2,+7 ; Quebec+Sherbrook as control\n")
road.main.res_35_lead2$ATT
cat("0,+7 ; Quebec only as control\n")
road.main.res_q_lead2$ATT
cat("-2,+7 ; Quebec only as control\n")
road.main.res_q_35_lead2$ATT
cat("-3,+0 ; Quebec+Sherbrook as control\n")
road.main.res_35_38_lead2$ATT
cat("0,+3 ; Quebec only as control\n")
road.main.res_q_38_lead2$ATT
cat("-3,+0 ; Quebec only as control\n")
road.main.res_q_35_38_lead2$ATT

cat("Penalized, 0,+7 ; Quebec+Sherbrook as control\n")
road.main.res1se_lead2$ATT
cat("Penalized, -2,+7 ; Quebec+Sherbrook as control\n")
road.main.res_351se_lead2$ATT
cat("Penalized, 0,+7 ; Quebec only as control\n")
road.main.res_q1se_lead2$ATT
cat("Penalized, -2,+7 ; Quebec only as control\n")
road.main.res_q_351se_lead2$ATT
cat("Penalized, -3,+0 ; Quebec+Sherbrook as control\n")
road.main.res_35_381se_lead2$ATT
cat("Penalized, 0,+3 ; Quebec only as control\n")
road.main.res_q_381se_lead2$ATT
cat("Penalized, -3,+0 ; Quebec only as control\n")
road.main.res_q_35_381se_lead2$ATT


cat("-2,+7 ; Quebec only as control\n")
road.main.res_q_35_lead2$ATT
```




```{r}
#| class-output: center-table
#| message: true
#| include: true
#| warning: true
#| error: true
#| label: synth3_lead2
#| eval: true
#| results: hold
#| fig.width: 12
#K= Number of non-overlapping hold-out blocks (“folds”) used for cross-fitting the weights and penalty. More folds ⇒ less bias but higher variance and longer run time.
#When countSynth tunes its ridge/elastic-net penalty (λ) by cross-validation, this flag tells it not to use the “one–standard-error” rule. If you set it to TRUE, the algorithm would pick the most parsimonious λ within 1 s.e. of the CV minimum.

#prepost_rmspe_cntsh(get_covbal_csnth(road.main.res), t_int=38)

result_objects3 <- list(
  list(obj = road.main.res_lead2, name = "road.main.res_lead2", desc = "0,+7 ; Quebec+Sherbrook as control"),
  list(obj = road.main.res_35_lead2, name = "road.main.res_35_lead2", desc = "-2,+7 ; Quebec+Sherbrook as control"),
  list(obj = road.main.res_q_lead2, name = "road.main.res_q_lead2", desc = "0,+7 ; Quebec only as control"),
  list(obj = road.main.res_q_35_lead2, name = "road.main.res_q_35_lead2", desc = "-2,+7 ; Quebec only as control (exRepeat)"),
  list(obj = road.main.res_35_38_lead2, name = "road.main.res_35_38_lead2", desc = "-3,+0 ; Quebec+Sherbrook as control"),
  list(obj = road.main.res_q_38_lead2, name = "road.main.res_q_38_lead2", desc = "0,+3 ; Quebec only as control"),
  list(obj = road.main.res_q_35_38_lead2, name = "road.main.res_q_35_38_lead2", desc = "-3,+0 ; Quebec only as control"),

  # Penalized Models
  list(obj = road.main.res1se_lead2, name = "road.main.res1se_lead2", desc = "Penalized, 0,+7 ; Quebec+Sherbrook as control"),
  list(obj = road.main.res_351se_lead2, name = "road.main.res_351se_lead2", desc = "Penalized, -2,+7 ; Quebec+Sherbrook as control"),
  list(obj = road.main.res_q1se_lead2, name = "road.main.res_q1se_lead2", desc = "Penalized, 0,+7 ; Quebec only as control"),
  list(obj = road.main.res_q_351se_lead2, name = "road.main.res_q_351se_lead2", desc = "Penalized, -2,+7 ; Quebec only as control (exRepeat)"),
  list(obj = road.main.res_35_381se_lead2, name = "road.main.res_35_381se_lead2", desc = "Penalized, -3,+0 ; Quebec+Sherbrook as control"),
  list(obj = road.main.res_q_381se_lead2, name = "road.main.res_q_381se_lead2", desc = "Penalized, 0,+3 ; Quebec only as control"),
  list(obj = road.main.res_q_35_381se_lead2, name = "road.main.res_q_35_381se_lead2", desc = "Penalized, -3,+0 ; Quebec only as control")
)

# --------------------------------------------------------------------------
# 3. Loop through objects, print ATT, and create/store plots
# --------------------------------------------------------------------------

# Initialize an empty list to store the generated ggplot objects.
plot_list3 <- list()

# This loop iterates over each item in the `result_objects` list.
for (item in result_objects3) {

  # --- Print the ATT Value ---
  cat("--------------------------------------------------\n")
  cat("Processing:", item$name, "\n")
  cat(item$desc, "\n")
  print(item$obj$ATT)
  cat("\n")

  # --- Generate and Store the Plot ---
  # Check if the object exists to prevent errors
  if (exists(item$name) && !is.null(get(item$name))) {

    # Call the plotting function which returns a ggplot object
    figcscm <- plot_synth_cscm(
      res_object = item$obj,
      t_int_man = 35,       # Manual intervention time
      max_time = 38,        # Time horizon for the plot
      y_lab = "Crash rate", # Y-axis label
      colors = c("SCM" = "gray60", "CSCM" = "red", "Obs" = "black")
    )

    # Add the title using ggplot2's labs() function and store it
    plot_list3[[item$name]] <- figcscm + labs(title = NULL) + theme(plot.title = element_text(size = 13))

  } else {
    cat("Warning: Object", item$name, "not found. Skipping plot generation.\n")
  }
}

# --------------------------------------------------------------------------
# 4. Arrange and display the plots in 2x2 grids using cowplot
# --------------------------------------------------------------------------

# Now, print the combined plots. In a Quarto/Rmd file, this will render
# the grids directly in the output document.
png("_figs/supp_mat_cntsynts_3min_0_lead2.png", width = 800*4, height = 600*4, res = 300)
plot_grid(
  plotlist = plot_list3[c(5, 7, 12, 14)],
  ncol      = 2,
  nrow      = 2,
  labels    = c("a)", "b)", "c)", "d)"),  # ← aquí van las etiquetas
  label_size      = 16,      # tamaño del texto
  label_fontface  = "bold",  # negrita (opcional)
  label_x = 0.02, label_y = 0.98,  # posición relativa (0-1) dentro del panel
  hjust    = 0,  vjust    = 1      # alineación izq-arriba
)
dev.off()
#road.main.res_35_38_lead2; road.main.res_q_35_38_lead2; road.main.res_35_381se_lead2; road.main.res_q_35_381se_lead2
plot_grid(
  plotlist = plot_list3[c(5, 7, 12, 14)],
  ncol      = 2,
  nrow      = 2,
  labels    = c("a)", "b)", "c)", "d)"),  # ← aquí van las etiquetas
  label_size      = 16,      # tamaño del texto
  label_fontface  = "bold",  # negrita (opcional)
  label_x = 0.02, label_y = 0.98,  # posición relativa (0-1) dentro del panel
  hjust    = 0,  vjust    = 1      # alineación izq-arriba
)

cbind.data.frame(
  pen= c(rep(c("No","Yes"),each=2)),
  mods= c(rep(c("Quebec+Sherbrooke","Quebec only"),2)),
  rbind.data.frame(result_objects3[[5]]$obj$ATT,result_objects3[[7]]$obj$ATT, result_objects3[[12]]$obj$ATT, result_objects3[[14]]$obj$ATT)
) |> knitr::kable("markdown", caption="RRs")
```


<br>

# Session info

```{r session-info, echo=T, error=T, message=TRUE, paged.print=TRUE,eval=T, results='hold'}
cat(paste0("R library: ", Sys.getenv("R_LIBS_USER")))
cat(paste0("Date: ",withr::with_locale(new = c('LC_TIME' = 'C'), code =Sys.time())))
cat(paste0("Editor context: ", getwd()))
cat("quarto version: "); system("quarto --version") 

quarto::quarto_version()

save.image("_data/step2.RData")
```

```{r session-info-r, echo=T, error=T, message=TRUE, paged.print=TRUE,eval=T, results='hold'}
#| class-output: center-table

sesion_info <- devtools::session_info()
dplyr::select(
  tibble::as_tibble(sesion_info$packages),
  c(package, loadedversion, source)
) |> 
 knitr::kable(caption = "R packages", format = "html",
      col.names = c("Row number", "Package", "Version"),
    row.names = FALSE,
      align = c("c", "l", "r")) |> 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"),font_size = 12)|> 
  kableExtra::scroll_box(width = "100%", height = "375px")  
```

```{r session-info-python, echo=T, error=T, message=TRUE, paged.print=TRUE,eval=T, results='hold'}
#| class-output: center-table

reticulate::py_list_packages()%>% 
 knitr::kable(caption = "Python packages", format = "html",
      col.names = c("Package", "Version", "Requirement"),
    row.names = FALSE,
      align = c("c", "l", "r", "r"))%>% 
  kableExtra::kable_styling(bootstrap_options = c("striped", "hover"),font_size = 12)|>
  kableExtra::scroll_box(width = "100%", height = "375px")  
```


# References